### 2018-06-09 Conference/AAS 2015 OD EKF/EKF_IOD.tex

\documentclass[10pt]{article}
\usepackage[letterpaper,text={6.5in,8.6in},centering]{geometry}
\usepackage{amssymb,amsmath,amsthm,times,graphicx,subfigure,tabularx,booktabs,colortbl,multirow,threeparttable}
\usepackage[small]{caption}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\ave}[1]{\mathrm{E}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathfrak{d}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}

\renewcommand{\r}{\mathbf{r}}
\renewcommand{\u}{\mathbf{u}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\e}{\mathbf{e}}

\newcommand{\bfi}{\bfseries\itshape\selectfont}

\renewcommand{\baselinestretch}{1.2}

\date{}

\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{remark}{Remark}

\renewcommand{\thesubsection}{\arabic{subsection}. }
\renewcommand{\thesubsubsection}{\arabic{subsection}.\arabic{subsubsection} }

%\graphicspath{{/Users/tylee@seas.gwu.edu/Documents/Research/Estimation/Simulation/GloUncPro/DistSph/Build/Products/Debug/Diff/}}

\begin{document}
\thispagestyle{empty}

\noindent{\large \bf Air Force Summer Faculty Fellowship Program:}\vspace*{0.1cm}\\
\noindent{\Large \bf Nonlinear Observability for Relative Orbits with Line-of-Sight Measurements}\vspace*{0.1cm}\\
\noindent{T. Alan Lovell and Taeyoung Lee}\\
\noindent{July 25, 2013}\\

\subsection{Numerical Examples}

\paragraph{Initial Condition}

Five cases of the IOD data are simulated.

\paragraph{Extended Kalman Filter}

The covariance of the process noise is parameterized as follows:
\begin{align*}
Q_k = 10^{q}\times  \mathrm{diag}[10^{-8}I_{3\times 3}, \,10^{-10}I_{3\times 3}],
\end{align*}
where the integer parameter $q$ is varied as $q=-1,\,0,\,1,\,2$.

The resulting estimation errors are summarized at the table at the next page. Overall, the filter performed relatively well: 21 cases converged out of 26 cases for a proper level of process noise.

The cases where filter diverted can be categorized as follows:
\begin{itemize}
\item \textbf{IOD} When the initial estimate is poor, more explicitly when $e_{mag}(0)>9$ or $e_{mag}<0.05$, the filter diverged. This corresponds to the following cases:\\
Case 2B, Case 4A (IOD1)
\item \textbf{HCW} When the nonlinear trajectory is too close to the linearized HCW model, the filter diverged. This corresponds to the following case:
Case 1
\item \textbf{Process noise} Filter diverged when the process noise is too low or too high. This corresponds to the following cases:
Case 3B (IOD1), Case 3C (IOD1), Case 4B (IOD1), Case 4C (IOD2)
\end{itemize}

The numerical results when $q=0$ are illustrated at Figures 1-13. From these figures, filter divergence can be determined as follows:
\begin{itemize}
\item \textbf{Too low $\lambda_{\max}{P}$ } The maximum eigenvalue of $P$ drops instantaneously and remain at a very low number. It represents that the filter is falsely too confident about a wrong estimate. This corresponds to the following cases:\\
Case 1, Case 2B, Case 3B (IOD1)
\item \textbf{Too high error residual} The residual error is too large. This corresponds to the following cases:
Case 3C (IOD1), Case 4A (IOD1)
\end{itemize}

\begin{center}
\begin{threeparttable}[h]
\caption{Error variables for EKF with varying $Q_k$}
\begin{tabularx}{0.7\textwidth}{>{\centering $}X<{$}*{2}{>{$}c<{$}}|*{5}{>{$}c<{$}}}\toprule
\multirow{2}{*}{Case} & \multirow{2}{*}{IOD} & \multirow{2}{*}{$e_{mag}(0)$} & \multicolumn{4}{c}{$\bar e_{mag}$} \\
& & & q=-1 & q=0 & q=1 & q=2 \\\midrule
0A & IOD 1 & 0.9997  & 0.1129   & 0.1405  & 0.2859  & 0.4726 \\
0A & IOD 2 & 1.0012  & 0.0400   & 0.1470  & 0.2559  & 0.5269 \\
0B & IOD 1 & 0.9767  & 0.0153   & 0.0553  & 0.2734  & 0.7942 \\
0B & IOD 2 & 1.0088  & 0.0330   & 0.0531  & 0.2286  & 0.9262 \\
0C & IOD 1 & 1.0152  & 0.0441   & 0.0286  & 0.1191  & 0.6596 \\
0C & IOD 2 & 1.0139  & 0.0292   & 0.0290  & 0.2174  & 0.8249 \\\midrule
\mathbf{1} & IOD 1 & 0.9999  & (4.9960)   & (20.6584)  & (52.8485)  & (159.8720) \\
\mathbf{1} & IOD 2 & 1.0000  & (12.0855)   & (16.9126)  & (54.0381)  & (168.9124) \\\midrule
2A & IOD 1 & 1.0550  & 0.1925   & 0.1262  & 0.1677  & 0.6115 \\
2A & IOD 2 & 1.0238  & 0.1444   & 0.0283  & 0.1975  & 0.5445 \\
2B & IOD 1 & \mathbf{0.0429}  & (0.9810)   & (0.9494)  & (0.9120)  & (0.7438) \\
2B & IOD 2 & \mathbf{0.0429}  & (0.9813)   & (0.9506)  & (0.9069)  & (0.7565) \\
2C & IOD 1 & 0.9781  & 0.0313   & 0.0205  & 0.0284  & 0.0792 \\
2C & IOD 2 & 0.9935  & 0.0221   & 0.0230  & 0.0257  & 0.0842 \\\midrule
3A & IOD 1 & 1.0191  & 0.0841   & 0.0180  & 0.1445  & 0.3300 \\
3A & IOD 2 & 0.8170  & 0.0864   & 0.0416  & 0.1936  & 0.2913 \\
3B & IOD 1 & 1.0612  & (0.9761)   & (0.9785)  & 0.0283  & 0.0985 \\
3B & IOD 2 & 0.7685  & 0.0128   & 0.0275  & 0.0206  & 0.0864 \\
3C & IOD 1 & 7.8962  & (1.5927)   & (2.2080)  & 0.8156  & 0.7897 \\
3C & IOD 2 & 6.4921  & 0.9330   & 0.7311  & 0.6269  & 0.6311 \\\midrule
4A & IOD 1 & \mathbf{9.7347}  & (6.2650)   & (6.1141)  & (8.8398)  & (10.0667) \\
4A & IOD 2 & 0.8425  & 0.0212   & 0.0506  & 0.0228  & 0.0268 \\
4B & IOD 1 & 1.9276  & (0.2523)   & 0.2016  & 0.1417  & 0.1669 \\
4B & IOD 2 & 0.9379  & 0.0068   & 0.0147  & 0.0483  & 0.0157 \\
4C & IOD 1 & 1.1161  & 0.0347   & 0.0713  & 0.0187  & 0.0352 \\
4C & IOD 2 & 0.9646  & 0.0185   & 0.0191  & 0.0094  & (0.9824) \\
\bottomrule
\end{tabularx}
{\small
\begin{tablenotes}
    \item[$(\cdot )$] Filter diverges. 
  \end{tablenotes}}
\end{threeparttable}
\end{center}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0A_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0A_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0A_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0A_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 0A (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0B_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0B_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0B_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0B_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 0B (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0C_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0C_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0C_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_0C_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 0C (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_1_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_1_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_1_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_1_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 1 (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2A_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2A_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2A_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2A_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 2A (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2B_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2B_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2B_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2B_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 2B (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2C_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2C_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2C_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_2C_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 2C (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3A_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3A_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3A_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3A_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 3A (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3B_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3B_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3B_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3B_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 3B (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3C_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3C_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3C_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_3C_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 3C (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4A_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4A_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4A_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4A_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 4A (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4B_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4B_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4B_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4B_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 4B (IOD1: blue, IOD2: black)}
\end{figure}

\clearpage\newpage
\begin{figure}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4C_2_Q0_e}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4C_2_Q0_xx_norm}}
}
\centerline{
\subfigure[Estimation error ($e_{mag},e_{dir}$)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4C_2_Q0_eigP}}
\hfill
\subfigure[Magnitude of state vector (true:solid, estimation:dashed)]{
	\includegraphics[width=0.44\textwidth]{EKF_IOD_4C_2_Q0_eres}}
}
\caption{Extended Kalman Filter: Case 4C (IOD1: blue, IOD2: black)}
\end{figure}

\end{document}