### 2018-06-09 Conference/SCITECH19/SCITECH19.tex

\documentclass[conf]{new-aiaa}
%\documentclass[journal]{new-aiaa} for journal papers
\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{longtable,tabularx}
\setlength\LTleft{0pt}

\usepackage{amssymb,url,times}%,subfigure}% amsthm is the one!
\usepackage{caption,subcaption,hyperref}
\usepackage{color,comment}

\newtheorem{prop}{Proposition}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\abs}[1]{\ensuremath{\left| #1 \right|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\ip}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Figure \ref{fig:#1}}
\newcommand{\tr}[1]{\mbox{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mbox{tr}\ensuremath{\!\bracket{#1}}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\renewcommand{\d}{\ensuremath{\mathbf{d}}}
\newcommand{\pair}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\!\langle #1 \rangle\!\rangle}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\ad}{\ensuremath{\mathrm{ad}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\graphicspath{{../../Fig/}}

\title{Autonomous Aerial for Topological Mapping of Mars Environments}

\author{Evan Kaufman\footnote{Graduate Research Assistant} and Taeyoung Lee\footnote{Associate Professor, AIAA Member Grade (if any)}}
\affil{Department of Mechanical and Aerospace Engineering, The George Washington University, Washington, DC, 20052}
% TODO: funding source

\begin{document}

\maketitle

\begin{abstract}
This paper proposes a 3D autonomous exploration approach designed for a flying vehicle while mapping the complex environment of the surface of Mars. Shannon's entropy serves as a measure of map uncertainty, where the robot goal is to minimize entropy, thereby maximizing map information. We propose a novel technique to predict future map entropy from depth sensors capable of scanning 3D spaces, and formulate autonomous exploration as an optimization problem based on map information gain and collision-free travel costs. The proposed approach is demonstrated with numerical simulations using contour data from the surface of Mars.
\end{abstract}

%\section{Nomenclature}
%
%{\renewcommand\arraystretch{1.0}
%\noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
%$A$  & amplitude of oscillation \\
%$a$ &    cylinder diameter \\
%$C_p$& pressure coefficient \\
%$Cx$ & force coefficient in the \textit{x} direction \\
%$Cy$ & force coefficient in the \textit{y} direction \\
%c   & chord \\
%d$t$ & time step \\
%$Fx$ & $X$ component of the resultant pressure force acting on the vehicle \\
%$Fy$ & $Y$ component of the resultant pressure force acting on the vehicle \\
%$f, g$   & generic functions \\
%$h$  & height \\
%$i$  & time index during navigation \\
%$j$  & waypoint index \\
%$K$  & trailing-edge (TE) nondimensional angular deflection rate
%\end{longtable*}}

\section{Introduction}

The topography of Mars has several benefits in future mission planning for landing spacecraft and  image distortion correction among other research objectives. The Mars Orbiting Laser Altimeter (MOLA) mission greatly improved the topography map of the surface, but was limited to an uncertainty of roughly eight meters~\cite{MOLA99}. The only precise topography knowledge about the surface comes from four rovers, two of which are inactive, and have covered only a small portion of the surface area of Mars.

Mapping and exploring Mars presents several challenges. Existing rovers are ground vehicles, incapable of traversing steep or dangerous terrain. Flying vehicles avoid this restriction, but special attention must be placed on their motion planning in this uncertain environment. Humans cannot directly give commands to the vehicles due to the signal time delay between Mars and Earth. Even if a humans are present on Mars, they can only provide sub-optimal motion commands on a small scale. These challenges motivate an autonomous approach to 3D mapping and exploration in uncertain and complex environments for a flying vehicle near the surface of Mars.

The study of 3D mapping and autonomous exploration has received substantial attention for applications on Earth. Occupancy grid mapping in 3D is a popular approach where an environment is decomposed into evenly-spaced cubic grid cells, where each cell is assumed as occupied or free~\cite{ThrBurFox05,WurHorBenStaBur10}. The goal is finding the probability of each cell being occupied. An exact solution of this probability was recently found by the authors using the stochastic properties of onboard sensors directly~\cite{KauLeeAiMos16,KauTakAiLee17}. Using an exact occupancy grid mapping algorithm in 3D, the topological information of the surface of Mars can be accurately computed in real-time.

Next, the robot must determine motions based on how much mapping knowledge it expects to gain, known as autonomous exploration. The most popular approach to this problem is known as frontier-based exploration~\cite{ZhuDinLinWu15,SenWan16,KleDor13}. The main idea is that the robot moves through previously-mapped free space toward the boundary between free space and uncertain space, known as a frontier. The robot takes measurements of this frontier, thereby pushing back the boundary. This process is repeated until the map is well-known. Despite the intuitive nature of this approach, there is no measure or prediction of how these robotic actions improve map information. Thus, frontier-based approaches rely on a heuristic rule, and are inherently suboptimal.

In contrast, entropy-based autonomous exploration approaches seek to minimize future map uncertainty. Shannon's entropy is a measure of map uncertainty; since minimizing future map uncertainty is equivalent to maximizing map information gain, these approaches select robotic actions to minimize entropy~\cite{StaGriBur05}. In~\cite{KauAiLee16}, an exact solution to expected entropy was developed, and in~\cite{KauTakAiLee18}, this concept is extended to 3D office-like environments assuming the environment can be approximated as vertically-uniform. Similarly, \cite{MauDakPet14} simplifies tasks by decomposing the 3D space into rooms. While these approaches can be effective for buildings, they are incompatible with the surface of Mars, which has complicated geological features that require careful consideration.

In this paper, we propose a 3D autonomous exploration algorithm that considers the 3D probabilistic occupancy grid map and sensor properties explicitly. The complicated and uncertain geometry of the surface of Mars is captured by aerial mapping because this approach can cover a large area effectively. An onboard depth sensor, such as a 3D laser scanner, gathers range measurements to determine the occupancy probabilities of 3D grid cells. Then, collision-free future pose candidates are considered. The predicted negative entropy changes from measurements passing through 3D space are maximized while also considering travel costs. The robot then uses Dijkstra's search in 3D for collision-free motion planning, and the complete process follows a receding-horizon framework for consistently using updated information. The proposed optimal approach is designed specifically for real-time exploration of the surface of Mars, where the flying robot mostly maps occupied space below the vehicle, but is also capable of looking ahead to capture steep and quickly-changing terrain. In short, the proposed autonomous exploration approach predicts future map uncertainty in 3D space to determine optimal actions that maximize map information gain while avoiding collisions.

The paper is organized as follows. The mapping probability and expected entropy calculation are defined in Section \ref{sec:ProbDef}. The proposed 3D exploration for information gain maximization is covered in Section \ref{sec:Explore3D}. Numerical simulations are shown in Section \ref{sec:NumExamples}, followed by conclusions.

\section{Problem Definition}
\label{sec:ProbDef}

In this section, we define the 3D occupancy grid map probability and expected map entropy from a future measurement.

\subsection{Probabilistic Occupancy Grid Mapping in 3D}

First we define the map, measurements, and poses as follows. Let 3D map $m$ be composed of $n_m$ evenly-spaced cubic grid cells with edge length $\alpha$. For index $i\in\braces{1,2,\ldots,n_m}$, let $\mathbf{m}_i$ and $\bar{\mathbf{m}}_i$ be the events that the $i$-th grid cell are occupied and free, respectively, where each cell occupancy is assumed static, binary, and independent of other cells. At the $t$-th time step, we assume a known current pose $X_t=\braces{x_t,R_t}$ such that $x_t\in\Re^3$ is a 3D location and $R_t\in\SO$ is the 3D attitude, and history of poses $X_{1:t-1}$ is assumed known as well. Considering that most modern sensors return scans with multiple measurement rays, let measurement ray $z\in\Sph^2$ belong the the most recent scan $Z_t$, and let $Z_{1:t-1}$ be the history of measurement scans.

Next, we present recent contributions to occupancy probability determination from~\cite{KauLeeAiMos16,KauTakAiLee17}, which uses a simplified result of a complicated Bayesian probability. Let a temporary reduced map $r\subset m$ be composed of $n_r$ cells along measurement $z$, indexed by increasing distance from the current pose $X_t$. The cells of $r$ are easily acquired through 3D ray casting: the measurement ray vector from $x_t$ intersects 3D grid cells, and the associated depths are calculated with Euclidean distance. Let $\mathbf{r}_{k+}$ be the event that the $k$-th cell of reduced map $r$ is the closest occupied space, and the \emph{forward sensor model} $p(z|\mathbf{r}_{k+},X_t)$ is known from the stochastic properties of the depth sensor (e.g. a Gaussian distribution). Then, the probability of $\mathbf{r}_k$ based on $z$ is commonly referred to as the \emph{inverse sensor model}, which can be written compactly as
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})&=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1}),
\end{align}
where $\eta$ serves as a normalizer,
\begin{align}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_i^-\bigg]^{-1},
\end{align}
and the probability before normalizing is
\begin{align}
\label{eqn:Unnormalized}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},&Z_{1:t-1})=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]
+ \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-,
\end{align}
where a priori probability is $\mathbf{P}_k^-=P(\mathbf{r}_{k}|X_{1:t-1},Z_{1:t-1})$, its complement is $\bar{\mathbf{P}}_k^-=1-\mathbf{P}_k^-$, $P(\bar{\mathbf{r}}_{0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ for convenience, and $p(z|\mathbf{r}_{(n_r+1)+},X_t)$ represents the forward sensor model of a maximum sensor reading. By using repeated terms in \refeqn{allEta} and \refeqn{Unnormalized}, this equation has linear complexity with $n_r$ for all cells in reduced map $r$, amortized to $\mathcal{O}(1)$ for each cell, which is easily computed in real-time.

In short, an exact solution to occupancy probability can be computed from the most recent measurement in real time provided the stochastic sensor properties and a priori probabilities. This solution avoids common approximations or machine learning, and simplifies future measurement predictions, described next with expected entropy calculation.

\subsection{Expected Measurement Ray Entropy}

Here we define entropy as a measure of 3D occupancy grid map uncertainty. Shannon's entropy~\cite{StaGriBur05} for the $i$-th cell and the complete map are
\begin{align}
\label{eqn:ShannonsEntropyCell}
H(P(\mathbf{m}_i))&=-P(\mathbf{m}_i)\log{P(\mathbf{m}_i})-P(\bar{\mathbf{m}}_i)\log{P(\bar{\mathbf{m}}_i}),
\\
\label{eqn:ShannonsEntropyMap}
H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i)),
\end{align}
respectively. This serves as an effective measure because \refeqn{ShannonsEntropyCell} is minimized for $P(\mathbf{m}_i)=0$ and $P(\mathbf{m}_i)=1$ (smallest uncertainty), but is maximized when $P(\mathbf{m}_i)=0.5$ (largest uncertainty).

Next, we define a candidate future pose. Let candidate location $x_c\in\Re^3$ be the 3D location of a future pose, and $R_c\in\SO$ be the candidate attitude, such that the candidate future pose is $X_c=\braces{x_c,R_c}$. Consider measurement $z_c$, which is a possible measurement from pose $X_c$. Since candidate location $x_c$ and the grid cell locations are known, the distance from the robot to the $k$-th cell, namely $z_{c,k}$ is easily solved with 3D ray casting.

Finally, we present the fundamental equations to predict map information gain from a potential future measurement based on~\cite{KauAiLee16}. The measurement ray $z_c$ can only change the cell entropies along its ray, and therefore the reduced map $r\subset m$ is used, similar to mapping. The history of poses and measurements are required for subsequent equations as they affect a priori probabilities, but are removed below for simplicity. The entropy after $z_c$ updates the map is predicted as
\begin{align}
\label{eqn:DiscExpEntropyRay}
&\text{E}[H(P(r|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(\mathbf{r}_k|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\}.
\end{align}
The first term of the summand, namely $H(P(\mathbf{r}_k|x_c,z_{c,k}))$, is simply the entropy \refeqn{ShannonsEntropyCell} of the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized}. The second term of the summand, namely $P(z_{c,k}|x_c)$, can be solved by combining \refeqn{allEta} as
\begin{align}
\label{eqn:ProbMeas}
P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}},
\end{align}
where $\eta_{c,k}$ refers to the normalizer based on the measurement $z_{c,k}$.

In short, the expected entropy along a ray is calculated by decomposing the measurement into possible range values with 3D ray casting. Then, we find the entropy from each range value, multiplied by the probability of that outcome using terms from the inverse sensor model. The computational complexity is $\mathcal O(n_r^2)$, but this can be reduced substantially by neglecting highly-improbable measurement outcomes, shown in~\cite{KauAiLee16}.

\section{Autonomous Exploration in Complex 3D Environments}
\label{sec:Explore3D}

In this section, we propose how multiple measurement rays in 3D can determine map information gain from a pose candidate. Then, we show how a collision-free trajectory is determined with a 3D Dijkstra's search. Finally, we combine 3D information gain and travel costs into a single optimization.

\subsection{Map Information Gain in 3D}

The expected entropy for an arbitrary 1D ray spanning 3D space can be calculated with \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas} in a computationally tractable manner. However, computing the expected entropy from multiple rays simultaneously has exponential complexity, and is therefore computationally intractable. Additionally, numerous rays of a single scan commonly intersect the same grid cells, making consideration of \emph{every} measurement ray unnecessary.

Instead of considering expected entropy from a complete scan, we propose a real-time solution that selects sample measurements to determine an optimal attitude at each candidate pose and the associated map information gain. We assume the vehicle is capable of level flight, so the third axes of the world and body are aligned (Fig. \ref{fig:transforms}), so $R_c$ can be expressed as
\begin{align*}
R_c=\begin{bmatrix}
\cos\psi & -\sin\psi & 0
\\
\sin\psi & \cos\psi & 0
\\
0 & 0 & 1
\end{bmatrix}, \quad 0\leq\psi<2\pi,
\end{align*}
where $\psi$ represents the angle about the third body-fixed axis. The direction of a 3D measurement may also have a nonzero component in the vertical direction. This is achieved by rotating the sensor frame angle $\theta$ about the second sensor-fixed axis,
\begin{align*}
R_{c,\text{sensor}}=\begin{bmatrix}
\cos\theta & 0 & \sin\theta
\\
0 & 1 & 0
\\
-\sin\theta & 0 & \cos\theta
\end{bmatrix}, \quad -\frac{\pi}{2}\leq\theta\leq\frac{\pi}{2},
\end{align*}
where $\Theta$ is the angular range for sensor field-of-view in the vertical direction. Combining these two rotation matrices, the measurement $z$ with depth $\norm{z}$ is expressed with respect to a frame fixed to Mars as
\begin{align}
z(\psi,\theta)=\begin{bmatrix}\cos\theta\cos\psi & -\sin\psi & \cos\psi\sin\theta
\\
\cos\theta\sin\psi & \cos\psi & \sin\psi\sin\theta
\\
-\sin\theta & 0 & \cos\theta
\end{bmatrix}
\begin{bmatrix}
\norm{z} \\ 0 \\ 0
\end{bmatrix}
=
\begin{bmatrix}
\cos\theta\cos\psi
\\
\cos\theta\sin\psi
\\
-\sin\theta
\end{bmatrix}\norm{z}.
\end{align}
In short, unit vectors are acquired via Euler angles within certain sensor limits.

% trim={<left> <lower> <right> <upper>}

\begin{figure}[!t]
	\centering
	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim = {21cm 5cm 20cm 5cm}, clip, height=1.0\textwidth]{mars_general.png}
        		\caption{General View}
    	\end{subfigure}
    	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim = {25cm 6cm 16cm 4cm}, clip, height=1.0\textwidth]{mars_psi.png}
        		\caption{$\psi$ Rotation}
    	\end{subfigure}
	\begin{subfigure}[t]{0.3\columnwidth}
           	\centering
          	\includegraphics[trim = {23cm 4cm 17cm 6cm}, clip, height=1.0\textwidth]{mars_theta.png}
        		\caption{$\theta$ Rotation}
    	\end{subfigure}
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{The three frames are shown with three axes (red: first axis, green: second axis, blue: third axis). The Mars frame (short and thick) is fixed to the planet. The body-fixed frame (medium thickness and length) is fixed to the camera frame (long and narrow).}
\label{fig:transforms}
\end{figure}

Next, we show how multiple 3D measurements provide an estimate for information gain. Let $\Psi$ and $\Theta$ be the angular ranges for sensor field-of-view (FOV) in the horizontal and vertical directions, respectively. Within the sensor FOV, we select sample measurements for evaluating their expected information gains. Let the number of sample measurements be $n_\psi$ and $n_\theta$ for horizontal and vertical rotations, respectively. Since $R_c$ and the FOV are known, the set of sample measurement rays, namely $\mathcal{Z}(R_c)$, is known as well. Then, the expected information gain for a single measurement ray $z$ from candidate position $x_c$ and attitude $R_c$ is the negative expected entropy change,
\begin{align}
\label{eqn:Iray}
\mathcal{I}_\text{ray}(x_c,R_c,\psi,\theta)=
\begin{cases}
H(P(r))-\text{E}[H(P(r|x_c,z(\psi+\tilde{\psi},\theta+\tilde{\theta})))], & \mbox{if } z(\psi+\tilde{\psi},\theta+\tilde{\theta})\in\mathcal{Z}(R_c), \\ 
0,                                                                                    & \mbox{otherwise},
\end{cases}
\end{align}
where the mounting yaw $\tilde{\psi}$ and pitch $\tilde{\theta}$ of the sensor is a fixed rotation in the horizontal and vertical directions, respectively in that order, such that $-\pi<\tilde{\psi}\leq\pi$ and $-\frac{\pi}{2}+\frac\Theta2\leq\tilde{\theta}\leq\frac{\pi}{2}-\frac\Theta2$. A positive $\tilde{\theta}$ corresponds to rotating the depth sensor from forward to downward toward the surface of Mars. The expected information gain for the full scan is the summation of expected information gains for individual rays
\begin{align}
\label{eqn:Iscan}
\mathcal{I}_\text{scan}(x_c,R_c)=\sum_{i_\psi=0}^{n_\psi-1}\sum_{i_\theta=0}^{n_\psi-1}\mathcal{I}_\text{ray}\left(x_c,R_c,\frac{\Psi i_\psi}{n_\psi},\frac{\Theta i_\theta}{n_\theta}\right).
\end{align}

Finally, \refeqn{Iscan} is computed for $n_\psi$ possible attitudes when the robot is at location $x_c$, one corresponding to each yaw angle $\psi$. Then, the optimal attitude at $x_c$ is
\begin{align}
\label{eqn:RcStar}
R_c^*=\argmax_{R_c} \mathcal{I}_\text{scan}(x_c,R_c).
\end{align}

In short, the expected information gain from each possible measurement ray is obtained using predicted entropy from \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas}. Then the expected information gain from all possible scans at a candidate location is calculated from \refeqn{Iray} and \refeqn{Iscan}. Finally, the optimal attitude at this candidate location is found using \refeqn{RcStar}.

\subsection{Collision-Free Trajectory in 3D}

Next, we present a fairly straightforward application of Dijkstra's search to 3D occupancy grid mapping. First we reduced the number of grid cells to cubic blocks large enough for a robot to fit completely inside. Based on the occupancy probability of these blocks and their neighbors, we determine which cells are reachable, and what are their travel costs by building a cost map based on Dijkstra's algorithm.

Here we describe how a reduced map is generated for collision avoidance and motion planning. If the largest edge-to-edge distance of the robot, namely $\rho$, is not larger than the 3D grid cell size $\alpha$, this step is unnecessary. Let $k\geq1$ be an integer such that $k\alpha\geq\rho$ and $(k-1)\alpha<\rho$, i.e., $k$ is the minimum number of grid cells in each dimension capable of fully-enclosing the robot inside this cube. Then, we decompose the complete probabilistic 3D into map $m_\text{reduced}$, which is composed of larger cubic cells that encompass $k^3$ cells from map $m$. Let $\mathbf{m}_{\text{reduced},k}$ denote the $k$-th cell of $m_\text{reduced}$ being occupied, and let $a_k\subset m$ be those $k^3$ cells from $m$ composing this larger cell. Then, we apply an effective rule for collision-avoidance from~\cite{KauTakAiLee18} to obtain the probability of $\mathbf{m}_{\text{reduced},k}$,

\begin{align}
\label{eqn:Proj3DMapComb}
P(\mathbf{m}_{\text{reduced},k})= 
\begin{cases}
    \max_{i\in a_k}{P(\mathbf{m}_i)},	&\text{if} \ \max_{i\in a_k}{P(\mathbf{m}_i)}\geq P_\text{thresh},\\
    \min_{i\in a_k}{P(\mathbf{m}_i)},	& \text{otherwise},
\end{cases}
\end{align}
where initial probability is $0<P_\text{init}<1$ and threshold the threshold probability is constrained to $P_\text{init}<P_\text{thresh}<1$. This approach uses probabilities that have changed from $P_\text{init}$, and favors occupied cells to avoid risking collisions.

Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search~\cite{Dij59}. Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$. A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidian distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away. Generating a cost map for all reachable cells in $\mathbf{m}_{\text{reduced},k}$ provides a collision-free travel cost for all reachable candidate pose locations. Once an optimal pose is selected, the path is easily found from the cost map using steepest descent.

\subsection{Optimal 3D Pose}

Here we present how an optimal pose is selected based on information gains at optimal attitudes and known collision-free travel costs. First, we show a function to account for travel costs, and then we account for these costs inside the optimization.

The proposed autonomous exploration follows a receding-horizon framework, meaning that optimal poses and trajectories are computed as quickly as possible, even if the robot has not reached the prior optimal pose. However, the computation is not instantaneous; let the distance $d_\text{opt}$ denote the distance that the robot can travel between optimal pose updates. Then, we define a \emph{bump function} loosely based on~\cite{Joh06}, defined as
\begin{align}
\label{eqn:BumpFunRecedingHorizon}
\mathcal B(d)=
\begin{cases}
f_\text{max},												& \text{if }0\leq d\leq d_\text{opt}
\\
(f_\text{max}-f_\text{far})\exp\braces{-\beta(d_\text{opt}-d)^2}+f_\text{far},	& \text{if }d>d_\text{opt}
\end{cases},
\end{align}
where $d$ is a travel distance along the cost map, $f_\text{max}>f_\text{far}>0$, and $\beta>0$ is an exponential decay rate of the bump, which is illustrated in Fig. \ref{fig:recedingHorizonBumpFun}. The main idea is that when travel costs are low, the bump function is large, but the bump decreases with increasing travel cost.

	\begin{figure}
	\vspace*{-0.3\columnwidth}
		\centerline{
			\includegraphics[width=0.6\columnwidth]{recedingHorizonBumpFunFlattened.pdf}
		}
	\vspace*{-0.25\columnwidth}
		\caption{The bump function is designed to prioritize local trajectories to avoid traversals across the map. However, there is no time wasted up to $d_\text{opt}$ because this distance may be achieved in the minimum exploration computation time.}
		\label{fig:recedingHorizonBumpFun}
	\end{figure}

Finally, we incorporate expected information gain and travel costs into a unified optimization. Given $d(x_c)$ from Dijkstra's cost map, the optimal candidate location is
\begin{align}
x_c^*=\argmax_{x_c}\mathcal{I}_\text{scan}(x_c,R_c^*)\mathcal B(d(x_c)),
\end{align}
and therefore the optimal pose is $X_c^*=\braces{x_c^*,R_c^*}$.

In conclusion, every reachable candidate pose is considered for its optimal attitude. The travel costs, acquired from Dijkstra's search, serve as input for the bump function, which prioritizes local trajectories to avoid costly map traversals. Then, the optimal candidate location is selected to maximize a combination of expected information gain and travel cost.

%\begin{itemize}
%	\item Describe the entropy calculation of expected information gain of all ray combinations is computationally intractable
%	\item Proposed strategy: multiple rays in 3D, and optimal attitude covers the rays with the greatest summation of information gains
%	\item Dijkstra's search in 3D, polynomial trajectories
%	\item Complete optimization
%	\item Concluding paragraph
%\end{itemize}

\section{Numerical Examples}
\label{sec:NumExamples}

\begin{itemize}
	\item Mars Environment: simulation environment generation, depth sensor facing slightly downward
	\item Software structure: ROS nodes, full map/reduced map for entropy options
	\item Numerical Results: reduced map used for entropy, full map used for entropy, Figs: RGB/heightmap/Gazebo, 3D colored grid cells with candidates (both full and reduced maps), entropy over time (full and reduced on same plot)
\end{itemize}
% software structure
% Numerical Result
  % reduced map used for entropy
  % full map used for entropy

\section{Conclusions}

\bibliography{../../BibSources}% master source for all publications

\end{document}