### 2018-06-09 Dissertation/Thesis/tex/ae2D.tex

% !TEX root = ../thesis-sample.tex

\chapter{Autonomous Exploration in 2D Space} \label{chap:ae2D}

\section{Entropy-Based Exploration}

\subsection{Shannon's Entropy}
% Description, figure

\subsection{Expected Information Gain}
% Derivation and link to Appendix

\subsection{Computational Limitations and Approximations}
% SIMPAR figure on entropy/computation time relationship

\section{Future Pose Optimization}

\subsection{Optimal Attitude from Sample Rays}
% SIMPAR figure on optimal pose

\subsection{Travel Cost with Obstacles}
% Dijkstra's cost map

\subsection{Optimal Pose Selection}
% Single-Vehicle argmax

\subsection{Collision-Free Motion Planning}
% Dijkstra's algorithm (after cost map), no least squares yet

\section{Numerical Examples}

\subsection{Exploring a Simple Environment}
% SIMPAR numerical example (2 rooms, 1 hallway)

\subsection{Exploring a Complicated Benchmark Environment}
% JINT17 Fig. 7 & 8

\section{Experimental Result}
% JINT17 9, 10, 11 (ground experiment at the NRL)

\section{Conclusions}