### 2018-06-09 Dissertation/Thesis/tex/ae3Dcomplex.tex

% !TEX root = ../thesis-sample.tex

\chapter{Autonomous Exploration in Complex 3D Space} \label{chap:ae3Dcomplex}

\section{Exploration with Full 3D Maps}

\subsection{Collision Mapping in 3D}

\subsection{Future Pose Optimization in 3D}

\section{Numerical Simulation}
% Mars simulation

\section{Experiment}
% SEH indoor simulation (needs completing)

\section{Conclusions}