### 2018-06-09 Dissertation/Thesis/tex/ae3Dsimple.tex

% !TEX root = ../thesis-sample.tex

\chapter{Autonomous Exploration in Vertically-Uniform 3D Space} \label{chap:ae3Dsimple}

\section{Occupancy Grid Mapping in 3D}

\section{Exploration in Simple Environments }

\subsection{Map Projections}

\section{Exploring a Large Room}
% IRC18 result: Fig. 4, 5, 6, 7

\section{Conclusions}