### 2018-06-09 Dissertation/Thesis/tex/multivehicle.tex

% !TEX root = ../thesis-sample.tex

\chapter{Multi-Vehicle Autonomous Exploration and Patrol} \label{chap:multivehicle}

\section{Bidding-Based Autonomous Exploration}

\subsection{Objective Function for the First Auction}

\subsection{Objective Function for Subsequent Auctions}

\subsection{Receding Horizon Framework}

\subsection{Multi-Vehicle Exploration Numerical Simulation}
% Multi-vehicle ICRA figs 3-6 (remake fig. 5 with larger contrast)

\section{Multi-Vehicle Cooperative Patrol}

\subsection{Continuous-Time Markov Process for Cell Degradation}
% JINT18 fig. 3

\subsection{Multi-Vehicle Patrol Numerical Simulation}
% JINT18 fig. 10-13 (ref. ICRA fig 3 from above in this chapter)

\section{Conclusions}