### 2018-06-20 Dissertation/Thesis/dissertation.tex
beginnings of mapping lit review

\documentclass[thesis]{thesis-gwu}

% this package is only used to generate some random text. 
% it is not needed in a true document
\usepackage{lipsum}

\graphicspath{{../../Fig/}}

\input{tex/frontmatter}

%% DOCUMENT AREA
\begin{document}

%\include{tex/intro}

\chapter{Introduction} \label{chap:intro}

This dissertation explores three key aspects of mobile robotics: occupancy grid mapping, autonomous exploration, and cooperation among robot teams. Several novel contributions are presented on both related topics for complex and dynamic situations.

\section{Motivation and Goals}

The study of robotics has expanded tremendously in recent years as robots have taken on dangerous, difficult, and repetitive tasks instead of humans. Certain tasks, such as search-and-rescue, surveillance, and cleaning, require some level of autonomy, particularly with mobile robots exploring new and uncertain environments. While these tasks provide significant value to our society, they present nontrivial challenges for the robots, particularly in mapping, exploration, and cooperation.

\subsection{Building Probabilistic Maps}

The first topic this dissertation addresses is building accurate probabilistic maps, known as \emph{mapping}. Map generation is crucial for simultaneous localization and mapping (SLAM) because maps provide important information for the robot to determine its pose (localization), which includes its position and attitude. Mapping also provides knowledge for trajectory planning through an environment and collision avoidance with mapped obstacles. In short, mapping is of fundamental importance in mobile robotics.

Therefore, mapping quality deserves much attention. Robotics is inherently \emph{probabilistic}; no sensor reading or action from an actuator is deterministic. All robotic processes have some degree of uncertainty, and the mapping must account for this. Onboard sensors serve to improve the understanding of surrounding spaces, but the stochastic properties of those sensors must be reflected in the stochastic properties of the map. Therefore, we propose to develop probabilistic map that accounts for the history of measurements and poses, and their associated stochastic properties. Here, each element of the map holds a probability based on measured data. In turn, this probability can be applied to tasks such as collision-avoidance and predicting future map outcomes.

\subsection{Exploration in Uncertain Spaces}

The second topic of this dissertation focusses on the motions robots must choose to maximize their knowledge about the map while avoiding collisions, known as \emph{autonomous exploration}.

\subsection{Cooperation Among Autonomous Systems}

The final topic of this dissertation is cooperation among members of a robot team.

\section{Literature Review}

In this section, we discuss the existing approaches to occupancy grid mapping, autonomous exploration, and multi-vehicle cooperation. Several shortcomings with existing approaches are highlighted, which are addressed in the dissertation contributions.

\subsection{Occupancy Grid Mapping}

Here we describe the existing approaches to building probabilistic occupancy grid maps. Though several variations have been proposed, no approach uses the stochastic properties of depth sensors directly to obtain an exact Bayesian solution that can be computed in real-time.

\paragraph{Inverse Sensor Model} The key to generating a probabilistic occupancy grid map is known as the \emph{inverse sensor model}~\cite{ThrBurFox05}. The main idea is that a depth sensor has a probability density (e.g. Gaussian) that describes the stochastic properties of a sensor reading (e.g. a LIDAR range) with respect to the expected value of a reading (e.g. distance to a wall). This probability density is referred to as the \emph{forward sensor model}, as it relates a depth measurement conditioned on the occupancy of the map.

In contrast, the inverse sensor model relates the occupancy of the map conditioned on the depth measurement. This problem is solved using Bayes' theorem. However, the occupancy grid map properties yield a complicated Bayesian probability; the computational cost has exponential complexity with respect to the number of grid cells captured within the depth sensor limits. It was previously-believed that this restriction makes solving the inverse sensor model in real-time impossible. This motivated approximate or estimated solutions, described next.

\paragraph{Approximated Function for the Inverse Sensor Model} Elfs et al originally proposed occupancy grid mapping in 1988...

% First proposed in 80s 
% Log-odds ratio solution
% simplified function - lines, patched Gaussians

% Learning
% Hill-climbing

In short, other existing solutions to occupancy probability are inexact, and involve a potentially-harmful log-odds ratio assumption. These shortcomings motivate an exact solution to occupancy grid mapping that can be computed in real-time.

%It was previously-believed that the exact solution to occupancy grid mapping was impossible in real-time.

\subsection{Autonomous Exploration}

\subsection{Multi-Vehicle Cooperation}

\section{Outline of Dissertation}

ADD ONCE COMPLETE

\section{Contributions}

There are three key contributions in this dissertation. The first two are novel solutions that find exact occupancy probabilities and predicted entropies, respectively. These contributions improve the quality and efficiency of autonomous mapping and exploration of uncertain environments for single-vehicle missions. The third contribution extends these concepts to multi-vehicle scenarios, where multiple members of a team work together to autonomously explore and patrol large environments.

\subsection{Summary of Contributions}

The first contribution is an exact Bayesian solution to occupancy grid mapping. The proposed approach considers that occupied spaces occlude measurement behind them, which allows certain mapping outcomes to be systematically grouped together. The exact solution to occupancy grid cell probability, which was previously-believed to have exponential complexity, is now possible to solve in real-time. Numerous numerical simulations and experiments demonstrate that the proposed mapping approach in effective and inexpensive in 2D and 3D scenarios.

The second contribution is a new approach to autonomous exploration, where the robot predicts future occupancy grid map uncertainty, using a measure known as Shannon's entropy. This approach computes the expected values of grid cell entropies from potential future measurement rays with a novel approach to predict future measurement probabilities. These predictions are used to evaluate the benefit of possible robot actions to maximize map information gain. Then, we formulate autonomous exploration as an optimization problem to maximize an objective function that includes map uncertainty and travel costs. The probabilistic map is used for collision-avoidance and motion planning. These processes are designed for real-time applications, where the robot makes decisions as it learns information about the surrounding environment.

The final contribution coordinates the mapping and exploration efforts of multiple robots together. The mapping process is multi-threaded such that multiple members update the same occupancy grid. The exploration algorithm follows a bidding-based structure. Robots compete for tasks in a series of auctions, where the winning bid of one auction affects the bids of subsequent auctions, thereby coordinating the robot efforts. This cooperative approach is computationally-efficient in large environments, and follows a receding-horizon framework, such that map expected information gains are updated as quickly as possible. The proposed coordination is extended to autonomous patrol, where occupancy grid map probabilities are degraded over time, which promotes the robots to revisit areas periodically.

\subsection{List of Publications}

The following is a list of publications. These include research included in this dissertation, as well as contributions in control, data association, and estimation.

% TODO: update this list

\begin{itemize}
	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2016, submitted.
	\item E. Kaufman, T. Lee, and Z. Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, San Francisco, December 2016, accepted.
	\item E. Kaufman, T. Lee, Z. Ai, and I. S. Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, Boston, July 2016, pp. 5709-5715.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{The Journal of the Astronautical Sciences}, 63(1): pp. 60-80, 2016, doi: 10.1007/s40295-015-0082-9.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Minimum Uncertainty JPDA Filters and Coalescence Avoidance for Multiple Object Tracking,'' \textit{The Journal of the Astronautical Sciences}, 63(4): pp. 308-334, 2016, doi: 10.1007/s40295-016-0092-2.
	\item T. Wu, E. Kaufman, and T. Lee, ``Globally Asymptotically Stable Attitude Observer on SO(3),'' \textit{Proceedings of the 54th IEEE Conference on Decision and Control}, pp. 2164-2168, Osaka, Japan, December 2015.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, Williamsburg, VA, Jan. 2015, AAS 15-451.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, VA, Jan. 2015, AAS 15-432.
	\item E. Kaufman, K. Caldwell, D. Lee, and T. Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, Mar. 2014, ASC 14-2527.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp. 2709-2714, June 2014.
	\vspace*{0.5cm}
\end{itemize}

\include{tex/pogm}
\include{tex/ae2D}
\include{tex/ae3Dsimple}
\include{tex/ae3Dcomplex}
\include{tex/multivehicle}

\bibliographystyle{plain}
\bibliography{../../BibSources.bib}

% appendices must appear after
\include{tex/appendix}
\end{document}