### 2018-06-23 Dissertation/Thesis/dissertation.tex
I think I deleted everyting

\documentclass[thesis]{thesis-gwu}

% this package is only used to generate some random text. 
% it is not needed in a true document
\usepackage{lipsum}

\graphicspath{{../../Fig/}}

\input{tex/frontmatter}

%% DOCUMENT AREA
\begin{document}

%\include{tex/intro}

\include{tex/intro}
\include{tex/pogm}
\include{tex/ae2D}
\include{tex/ae3Dsimple}
\include{tex/ae3Dcomplex}
\include{tex/multivehicle}

\bibliographystyle{plain}
\bibliography{../../BibSources.bib}

% appendices must appear after
\include{tex/appendix}
\end{document}