### 2018-06-23 Dissertation/Thesis/tex/intro.tex

% !TEX root = ../thesis-sample.tex

\chapter{Probabilistic Occupancy Grid Mapping} \label{chap:pogm}

% Add here

\section{Inverse Sensor Model}

\subsection{Bayesian Probability}

\subsection{Computationally-Efficient Solution}
% Include Fig 1 from JINT 2017

\section{Multi-Dimensional Maps}

\subsection{Ray Casting}

\subsection{Combining Measurements from a Single Scan}

\paragraph{Ray-By-Ray Approach}

\paragraph{Synergistic Update Approach}

\subsection{Multi-Sensor Fusion}

\subsection{Practical Implications}

\section{Numerical and Experimental Examples}

\subsection{Mapping a Room}
% ACC16 comparison of exact/approx. inverse sensor models

\subsection{Kinect Measurement Scan Analysis}
% ACC16 Single scan exact/approx. comparison

\section{Conclusions}