### 2018-06-23 Dissertation/Thesis/tex/pogm.tex
started chapter 2: mapping

% !TEX root = ../thesis-sample.tex

\chapter{Probabilistic Occupancy Grid Mapping} \label{chap:pogm}

In this chapter, we present the exact solution to occupancy grid map probability. The Bayesian approach uses the stochastic properties of sensors, and exploits important patterns in conditional probabilities. This solution is simulated in several examples to show the efficacy of the approach.

\section{Mapping Problem Definition}

\section{Inverse Sensor Model}

\subsection{Bayesian Probability}

\subsection{Computationally-Efficient Solution}
% Include Fig 1 from JINT 2017

\section{Multi-Dimensional Maps}

\subsection{Ray Casting}

\subsection{Combining Measurements from a Single Scan}

Next, we describe two approaches to combine the solutions to the inverse sensor model for a single measurement ray...

\paragraph{Ray-By-Ray Approach}

\paragraph{Synergistic Update Approach}

\subsection{Multi-Sensor Fusion}

\subsection{Practical Implications}

\section{Numerical and Experimental Examples}

\subsection{Mapping a Room}
% ACC16 comparison of exact/approx. inverse sensor models

\subsection{Kinect Measurement Scan Analysis}
% ACC16 Single scan exact/approx. comparison

\section{Conclusions}