### 2018-06-23 Dissertation/Thesis/tex/pogm.tex

% !TEX root = ../thesis-sample.tex

\chapter{Probabilistic Occupancy Grid Mapping} \label{chap:pogm}

In this chapter, we present the exact solution to occupancy grid map probability. The Bayesian approach uses the stochastic properties of sensors, and exploits important patterns in conditional probabilities. This solution is simulated in several examples to show the efficacy of the approach.

\section{Mapping Problem Definition}

Let a map $m$ be decomposed into $n_m$ evenly-spaced grid cells, where the $i$-th grid cell is assigned to a static binary random variable $\mathbf{m}_i$ for $i\in\braces{1,2,\ldots,n_m}$, that is defined as $\mathbf{m}_i=1$ when occupied, and $\mathbf{m}_i=0$ when free. The location and size of each grid cell is assumed known, where a smaller cell size (greater grid resolution) better represents a space, but increases computation and memory. Therefore, a map $m$ is defined by $\{\mathbf{m}_1,\ldots, \mathbf{m}_{n_m}\}$ ($2^{n_{m}}$ possible maps).

Another random variable is defined as $\bar{\mathbf{m}}_i=1-\mathbf{m}_i$ for convenience. The probability that the $i$-th cell is occupied is $P(\mathbf{m}_i)$, and the probability that it is free is $P(\bar{\mathbf{m}}_i)=1-P(\mathbf{m}_i)$. The random variables $\mathbf{m}_i$ are mutually independent, 
\begin{align}
P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i).
\end{align}

Consider a range sensor that provides scans of the surrounding environment in order to identify the closest occupied space. The location and the direction of the sensor, referred to as the \emph{pose} at time $t$, is denoted by $X_t=(x_t,R_t)$, where the planar position and direction of the robot at $t$ are denoted by $x_t\in\Re^2$ and $R_t\in\Sph^1=\braces{q\in\Re^2|\norm{q}=1}$, i.e., the attitude corresponds to a direction on a 2D plane.  Let $X_{1:t}$ denote the history of poses from the initial time to the current time, i.e., $X_{1:t}=\{X_1,X_2,\ldots, X_t\}$. At each pose, the robot receives a 2D measurement \emph{scan}, composed of $n_z$ measurement \emph{rays}. These rays are 1D depth measurements of known direction from the current pose to the closest occupied space, subject to a known \emph{forward sensor model} $p(z_{t,l}|m,X_t)$, where $z_{t,l}$ is the $l$-th measurement ray of the $t$-th scan $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$. The measurement history is denoted by $Z_{1:t}=\braces{Z_1,Z_2,\ldots,Z_t}$.

The probability density function, namely $p(z_{t,l}|m,X_t)$ with respect to the depth of the $l$-th measurement ray conditioned on the map $m$ and the pose $X_t$ is commonly referred to as the \emph{forward sensor model}, which characterizes the corresponding depth sensor, such as the maximum range or accuracy. The forward sensor model satisfies (i) the ranges of all depth measurements are positive and finite, and (ii) measurement rays cannot pass through occupied regions. Throughout this paper, we assume that the forward sensor model of the selected sensor is given. This can be determined empirically or analytically. For example, the \emph{beam model for range finders} satisfying  the above criteria is described in~\cite{ThrBurFox05}.

Occupancy grid mapping provides occupancy probabilities based on robot poses and measurement scans. The goal is to obtain $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$, commonly referred to as the \emph{inverse sensor model} for given forward sensor model $p(z_{t,l}|m,X_t)$ and the initial estimate of the map $P(m)$ with $X_{1:t}$ and $Z_{1:t}$.

\section{Inverse Sensor Model}

In this section, we propose an algorithm to compute the exact inverse sensor model efficiently. 
Suppose the probability of the map conditioned on the past poses and measurements, namely $P(m|X_{1:t-1},Z_{1:t-1})$, is known. Here we construct a posteriori probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$, based on the current pose $X_t$, the measurement from the $l$-th ray $z_{t,l}$, and the given forward sensor model $p(z_{t_l}|m,X_t)$.

\subsection{Bayesian Framework}

The occupancy probability $P(m|z_{t,l},X_{1:t},Z_{1:t-1})$ is based on the forward sensor model $p(z_{t,l}|m,X_{1:t},Z_{1:t-1})$ that describe the distribution of the measurements for the given robot pose and the map. This specifies the stochastic characteristics of the sensor, such as the mean squared error, and it is assumed as a known distribution, specific to a particular depth sensor. One could certainly obtain a stochastic model by fitting a probability density to a controlled set of measurements. Bayes' rule yields
\begin{align}
\label{eqn:BayesRuleRayISM}
P(&m|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
=\frac{p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1})}{p(z_{t,l}|X_{1:t},Z_{1:t-1})},
\end{align}
considering that $X_t$ carries no information about $m$ without $Z_t$.
If the current pose $X_t$ and map $m$ are known, then the measurement ray $z_{t,l}$ is independent of past poses $X_{1:t-1}$ and past measurements $Z_{1:t-1}$ to obtain
\begin{align}
P(m|z_{t,l},&X_{1:t},Z_{1:t-1})%\nonumber\\&
=\eta_{t,l}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}),
\end{align}
where the normalizing constant $\eta_{t,l}\in\Re$ absorbs all terms that do not depend on the map $m$.
Next, we compute the occupancy probability of each cell. Let $\mathcal{M}_i$ be the set of maps where the $i$-th cell is occupied, i.e., $\mathcal{M}_i =\{m\in\{0,1\}^{{n_m}}\,|\ \mathbf{m}_i=1\}$. To compute the probability of occupancy of the $i$-th cell, all possible combinations of map in $\mathcal{M}_i$ should be considered, i.e., 
\begin{align}
\label{eqn:InvSenModWithProbDens}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align}
Furthermore, to determine the normalizing constant $\eta_{t,l}$, the complement $\\P(\bar{\mathbf{m}}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ must be calculated in the similar manner. Since there are $2^{n_{m}-1}$ maps in $\mathcal{M}_i$, these equations require $2^{n_m}$ terms to compute the summation for the $i$-th grid cell and normalizer, and this process should be repeated for all other cells along the measurement ray. This is the main reason why the existing results are based on certain approximations of the above expression.

\subsection{Computationally-Efficient Solution}
% Include Fig 1 from JINT 2017

We propose a computational algorithm to evaluate \refeqn{InvSenModWithProbDens} efficiently. 
Since the cells outside of the sensor field of view (FOV) are not affected, we focus on a reduced map $r_l$ in the FOV of the $l$-th ray. This reduced map is chosen such that each cell of $r_l$ corresponds to a grid cell of map $m$ that the $l$-th ray intersects, ordered by increasing distance, which is easily determined from geometry. Let $\mathbf{r}_{l,k}$ be the binary random variable representing the occupancy of the $k$-th cell of the $l$-th ray. The number of cells in the reduced map is denoted by $n_{r,l}\leq n_m$.

Next, let $\mathbf{r}_{l,k+}$ correspond to the event that the $k$-th cell of the $l$-th ray is occupied, cells with lower index (closer cells) are free, and cells with greater index (farther cells) may or may not be occupied, i.e., event $\mathbf{r}_{l,k+}$ occurs when \\$r\in\braces{r\in\{0,1\}^{{n_{r,l}}}\,|\mathbf{r}_{l,1}=0,\mathbf{r}_{l,2}=0,\ldots,\mathbf{r}_{l,k-1}=0,\mathbf{r}_{l,k}=1}$. In other words, the $k$-th cell $\mathbf{r}_{l,k}$ is the closest occupied cell to current pose $X_t$ along the $l$-th ray. For example, see Figure \ref{fig:show_rkplus} that illustrates the event of $\mathbf{r}_{l,k+}$ when $l=1$ and $k=4$ for an one-dimensional cell array. This concept of grouping map outcomes can be easily extended to 2D using ray casting, where a 1D ray is spanning 2D space. The cells considered for the measurement ray are determined with the geometric intersection between the ray and map cell edges.
Then, the forward sensor model is identical for all maps defined by $\mathbf{r}_{l,k+}$, regardless of the occupancy of the cells beyond the $k$-th cell, and the corresponding forward sensor model $p(z_{t,l}|\mathbf{r}_{l,k+},X_{t})$ depends on the distance from $X_t$ to the $k$-th cell. Based on this, we present the mathematical expressions for the exact inverse sensor model, as summarized by \refeqn{RayISMAnswer} in Proposition 1. Later in Section \ref{sec:SA}, these are also rearranged as a computational algorithm.

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{rkplus_1.png}
    \centering
  \includegraphics[width=0.4\textwidth]{rkplus_2.png}  
  \centering
  \includegraphics[width=0.4\textwidth]{rkplus_3.png}
    \centering
  \includegraphics[width=0.4\textwidth]{rkplus_4.png}
  \caption{A robot located to the left of a 1D occupancy grid map $r_l$ composed of $n_{r,l}=6$ grid cells in $4$ cases. In each case, cells $1$--$3$ are free, cell $4$ is occupied, and cells $5$ and $6$ may or may not be occupied. All above outcomes correspond to the event $\mathbf{r}_{l,4+}$.}
  \label{fig:show_rkplus}
\end{figure}

\begin{prop}
For the $l$-th measurement ray, the a posteriori probability of the occupancy of the $k$-th cell, namely the ray inverse sensor model, is given by
\begin{align}
\label{eqn:RayISMAnswer}
P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})&=\eta_{t,l}\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1}),
\end{align}
where the unnormalized probability of the inverse sensor model is defined as
\begin{align}
\label{eqn:Unnormalized}
& \tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})%\nonumber\\&
=P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})\nonumber\\
&\quad\times 
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
%&\quad\times 
p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]\nonumber\\
&\quad + \bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}%\nonumber\\
%&\quad\times 
p(z_{t,l}|\mathbf{r}_{l,k+},X_t)P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1}),
\end{align}
where $P(\bar{\mathbf{r}}_{l,0}|X_{1:t-1},Z_{1:t-1})=P(\mathbf{r}_{l,n_r+1}|X_{1:t-1},Z_{1:t-1})=1$ is chosen for convenience and $p(z_{t,l}|\mathbf{r}_{l,(n_r+1)+},X_t)$ represents the probability density of the measurement when all of cells in the field of view of the $l$-th ray is not occupied. The normalizer $\eta_{t,l}$ is given by
\begin{align}
\label{eqn:allEta}
\eta_{t,l}
&=
\bigg[\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}\nonumber\\&\quad\times p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})\bigg]^{-1},
\end{align}
and it is independent of the cell index $k$.
\end{prop}
\begin{proof}
See Appendix.
\end{proof}

Note that the a priori estimate, $P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ and its compliment $\\P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})=1-P(\mathbf{r}_{l,k}|X_{1:t-1},Z_{1:t-1})$ are available at the $t$-th step. Then, \refeqn{RayISMAnswer} yields a sequential occupancy grid mapping that can be applied whenever new measurements are available.

% TODO: add update paragraph from JINT18

%Compared with \refeqn{InvSenModWithProbDens}, where the terms of the summation should be repeated $2^{n_{r,l}}$ times \emph{per each cell of the reduced map}, the proposed expressions \refeqn{RayISMAnswer} and \refeqn{allEta} are \textit{substantially} simpler. In fact, if \refeqn{Unnormalized} and \refeqn{allEta} are obtained recursively, requiring the summation of $O(n_{r,l}+1)$ rather than $O(n_{r,l}\times2^{n_{r,l}})$ as previously thought for \emph{all cells of the reduced map}, this yields an algorithm that is $\mathbf{n_{r,l}\times2^{n_{r,l}}/(n_{r,l}+1)}$ \textbf{times faster}.

\section{Multi-Dimensional Maps}

The above formulations describe how a single ray updates grid cells along its 1D path. Next we describe how measurement rays can update 2D maps, and how large scans of measurement rays are handled together.

\subsection{Ray Casting}

\subsection{Combining Measurements from a Single Scan}

Next, we describe two approaches to combine the solutions to the inverse sensor model for a single measurement ray...

\paragraph{Ray-By-Ray Approach}

\paragraph{Synergistic Update Approach}

\subsection{Multi-Sensor Fusion}

\subsection{Practical Implications}

\section{Numerical and Experimental Examples}

\subsection{Mapping a Room}
% ACC16 comparison of exact/approx. inverse sensor models

\subsection{Kinect Measurement Scan Analysis}
% ACC16 Single scan exact/approx. comparison

\section{Conclusions}