### 2018-06-24 Dissertation/Thesis/tex/ae2D.tex
added text from JINT17, needs revision

% !TEX root = ../thesis-sample.tex

\chapter{Autonomous Exploration in 2D Space} \label{chap:ae2D}

The goal of autonomous exploration is to select robotic actions designed to minimize map uncertainty, thereby maximizing map information gain. We formulate autonomous exploration as an optimization problem to determine a policy that accounts for map uncertainty and travel cost.

\section{Entropy-Based Exploration}

In this section, we define Shannon's entropy as a measure of map uncertainty, and present a novel approach to predict future entropy from an arbitrary ray. Then we discuss computational modifications for real-time implementation.

\subsection{Shannon's Entropy}

Here we present how the probabilistic properties of an occupancy grid map provide key information about the uncertainty of the space for motion planning. Shannon's entropy commonly serves as an uncertainty measure~\cite{StaGriBur05}. Provided grid cell probabilities of map $m$, Shannon's entropy is defined as
\begin{align}
\label{eqn:ShannonsEntropyCell}
H(P(\mathbf{m}_i))&=-P(\mathbf{m}_i)\log{P(\mathbf{m}_i})-P(\bar{\mathbf{m}}_i)\log{P(\bar{\mathbf{m}}_i}),
\\
\label{eqn:ShannonsEntropyMap}
H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i)),
\end{align}
for an individual cell and the entire map, respectively.
Thus, entropy is maximized when $P(\mathbf{m}_i)=0.5$, which corresponds to the largest uncertainty; similarly, entropy is minimized as $P(\mathbf{m}_i)$ approaches $0$ or $1$, which corresponds to the smallest uncertainty. %; thus Shannon's entropy is a measure of uncertainty.

\subsection{Expected Information Gain}

Suppose a future pose candidate and its associated future measurement scan is $X_c=\braces{x_c,R_c}$ and $Z_c$, respectively, where $c\in\mathcal C$ such that $\mathcal C=\braces{1,2,...,n_c}$ accounts for all $n_c$ candidates under consideration for the next robot pose. The current pose of the robot is \emph{not} $X_c$ in general, so any change to the probabilistic map from $X_c$ must be predicted. Much like the probabilistic mapping, this is achieved ray-by-ray~\cite{KauAiLee16,KauTakAiLee17}. Considering that all grid cell probabilities are conditioned on the history of poses $X_{1:t}$ and measurement scans $Z_{1:t}$, these terms are removed from the remaining equations for simplicity.

\begin{prop}
\label{prop:ExpectedH}
For candidate ray $z_c$, the expected entropy is
\begin{align}
\label{eqn:DiscExpEntropyRay}
&\text{E}[H(P(m|x_c,z_{c}))]=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\},
\end{align}
where $z_{c,k}$ refers to the distance from $x_c$ to the $k$-th grid cell along the measurement ray. The first term of the summation of \refeqn{DiscExpEntropyRay}, namely $H(P(m|x_c,z_{c,k}))$, is obtained with entropy definitions \refeqn{ShannonsEntropyCell}, \refeqn{ShannonsEntropyMap} and the inverse sensor model \refeqn{RayISMAnswer}--\refeqn{Unnormalized}. The second term is derived from \refeqn{allEta} with
\begin{align}
\label{eqn:ProbMeas}
P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}},
\end{align}
where $\eta_{c,k}$ refers to the normalizer based on the measurement $z_{c,k}$.
The expected negative entropy change for candidate pose $X_c$ is equivalently the expected information gain from ray $z_c$,
\begin{align}
\label{eqn:expectedInfoGainRay}
\mathcal I(X_c,z_c)&=H(P(m))-\text{E}\left[H(P(m|X_c,z_c))\right].
\end{align}
\end{prop}
\begin{proof}% TODO: add ref?
See Appendix B.
\end{proof}

The proposed approach discretizes the measurement according to assumptions of occupancy grid mapping, and exploits the probabilistic properties uncovered by the proposed inverse sensor model. These are used to directly calculate the expected value of map entropy for a single measurement ray.

\subsection{Computational Limitations and Approximations}

The computational order for each measurement ray is $\mathcal O(n_{r}^2)$ since the summations of \refeqn{ProbMeas} are embedded in \refeqn{DiscExpEntropyRay}. However, several of those intersections provide negligible information since the probability of the measurement ray capturing certain cell depths is close to zero.

The approximation of expected ray entropy provides a method to reduce the computation of \refeqn{expectedInfoGainRay} substantially. This goal is achieved by systematically selecting a smaller set of grid cells to consider over the summations of \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas}.
The smaller set is determined by the probability that each cell is captured by the measurement ray, known as the detection probability. This can be found recursively as
\begin{align}
\label{eqn:ProbOfFirstCell}
P(\mathbf{r}_{k+})%\nonumber\\&
=\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{j})\bigg\}P(\mathbf{r}_{k}),
\end{align}
which is the probability that $\mathbf{r}_{k}$ is the closest occupied grid cell based on past poses and measurement scans, independent of cells beyond the $k$-th cell from $x_c$.
Let $\hat n>0$ be a fixed number of grid cells for all rays such that $\hat n\leq n_{r}+1$ for all $l$.
Let $\hat{r}$ correspond to the grid cells that yield the $\hat{n}$ maximum values of \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.
By replacing the reduced map $r$ with $\hat{r}$ and changing the summation limits to $\braces{1,2,...,\hat n}$ in \refeqn{DiscExpEntropyRay} and \refeqn{ProbMeas}, the order of computation is reduced to $\mathcal O({\hat{n}}^2)$.
Even though the value of $n_{r}$ is different among various rays in general, $\hat n$ is fixed, so the computational order is fixed as well.
In short, this method reduces the required computation substantially by systematically neglecting those grid cells with little effect. It can be noted that if $\hat n=n_{r}+1$, the ray objective function is computed without approximation.

\paragraph{Algorithm} We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Algorithm \ref{alg:RayExpectedEntropyGain}). Much like the algorithm pseudocode for the ray-by-ray inverse sensor model (Algorithm \ref{alg:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification.

\vspace*{0.05\columnwidth}
\begin{algorithm}[H]
	Function: $\mathcal I_\text{ray}=\text{RayExpInfoGain}(x,P(r),z_{1:n_{r}})$\;
	Initialize $P(\bar{\mathbf{r}}_{0})=P(\hat{\bar{\mathbf{r}}}_{0})=P(\bar{\mathbf{r}}_{n_{r}+1})=1$\;
	\For{$k = 1,2,\ldots,n_{r}+1$}{
		$P(\mathbf{r}_{k+}) = P(\bar{\mathbf{r}}_{0:k-1})P(\mathbf{r}_{k})$\;
		$P(\bar{\mathbf{r}}_{0:k})=P(\bar{\mathbf{r}}_{0:k-1})(1-P(\mathbf{r}_{k}))$\;
	}
	Find $\hat r\subset r$ of the $\hat n$ greatest values of 
	$\braces{P(\mathbf{r}_{1+}),P(\mathbf{r}_{2+}),\ldots,P(\mathbf{r}_{(n_{r}+1)+})}$\;
	\For{$k = 1,2,\ldots,\hat{n}$}{
		$P(\hat{\mathbf{r}}_{k+}) = P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\mathbf{r}}_{k})$\;
		$P(\hat{\bar{\mathbf{r}}}_{0:k})=P(\hat{\bar{\mathbf{r}}}_{0:k-1})P(\hat{\bar{\mathbf{r}}}_{k})$\;
	}
	\For{$k_\text{m}=1,2,\ldots,\hat n$}{
		Initialize $\eta^{-1}_{k_\text{m}}=0$\;
		\For{$k_\text{c}=1,2,\ldots,\hat n$}{
			$a_\text{temp}=P(\hat{\mathbf{r}}_{k_\text{c}+})p(z_{k_\text{m}}|\hat{\mathbf{r}}_{k_\text{c}+},x)$\;
			$\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=P(\hat{\mathbf{r}}_{k_\text{c}})\eta^{-1}_{k_\text{m}}+a_\text{temp}$\;
			$\eta^{-1}_{k_\text{m}}=\eta^{-1}_{k_\text{m}}+a_\text{temp}$\;
		}
		$P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})=\eta_{k_\text{m}}\tilde P(\hat{\mathbf{r}}_{k_\text{c}}|x,z_{k_\text{m}})$ for all $k_\text{c}=1,2,\ldots,\hat n$\;
	}
	$P(z_{k_\text{m}}|x)=\frac{\eta^{-1}_{k_\text{m}}}{\sum_{i=1}^{\hat n}\eta^{-1}_{i}}$ for $k_\text{m}=1,2,\ldots,\hat n$\;
	Initialize $\mathcal I_\text{ray}=0$\;
	\For{$k_\text{c}=1,2,\ldots,\hat n$}{
		$\mathcal I_\text{ray}=\mathcal I_\text{ray}+H(P(\hat{\mathbf{r}}_{k_\text{c}}))$\;
		\For{$k_\text{m}=1,2,\ldots,\hat n$}{
			$\mathcal I_\text{ray} = \mathcal I_\text{ray}-H(P(\hat{\mathbf{r}}_{k_\text{c}}|x_c,z_{k_\text{m}}))P(z_{k_\text{m}}|x_{c})$\;
		}
	}
	Return: $\mathcal I_\text{ray}$\\
\caption{Expected Information Gain from a Measurement Ray}
\label{alg:RayExpectedEntropyGain}
\end{algorithm}

\paragraph{Numerical Justification for the Approximation}

The purpose of this numerical example is to provide evidence that the approximations are reasonable and increase the algorithm speed substantially.
Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have spacing $\alpha=0.2$ m, and the properties of the range sensor are based on the Microsoft Kinect~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$ m ($20$ grid cells inside the sensor FOV). The goal is to compare the expected entropy $\text{E}[H(P(m|x_c,z_{c}))]$ from \refeqn{DiscExpEntropyRay} and with an approximation $\text{E}[H_\text{approx}(P(m|x_c,z_{c}))]$, which only considers $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.

We consider $100$ probabilistic maps to obtain Monte Carlo results to evaluate the approximate entropy. In every Monte Carlo trial, each grid cell has an $80\%$ chance of being free and a $20\%$ chance of receiving an a priori probability uniformly distributed between $0$ and $1$. 
Several metrics serve to evaluate $\text{E}[H(P(m|x_c,z_{c}))]$ with $\text{E}[H_\text{approx}(P(m|x_c,z_{c}))]$. The median expected entropy change is $\text{E}[H(P(m|x_c,z_{c}))]-H(P(m|X_{1:t},Z_{1:t}))=-0.83792$. The error for the $100$ Monte Carlo cases is defined simply as
\begin{align}
e_{H}&=\frac1{100}\sum_{i=1}^{100}\text{abs}\bigg(\text{E}[H(P(m|x_c,z_{c}))]-\text{E}[H_\text{approx}(P(m|x_c,z_{c}))]\bigg).
\end{align}
The Monte Carlo trials are repeated for $\hat n=\braces{1,2,\ldots,10}$ and the results are plotted in Figure \ref{fig:ApproxJust}.
This example shows a typical case when the summation limits generated from \refeqn{ProbOfFirstCell} have only small effects on \refeqn{DiscExpEntropyRay}, while providing very large improvements in reducing computation.

\begin{figure}
	\centering
    	\begin{subfigure}[b]{0.45\textwidth}
        		\includegraphics[width=\textwidth]{JustifyApprox_eH.pdf}
        		\caption{Entropy Error}
        		\label{fig:H_err}
    	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
        		\includegraphics[width=\textwidth]{JustifyApprox_t.pdf}
        		\caption{Computation Ratio}
        		\label{fig:H_comp_ratio}
    	\end{subfigure}
\caption{The entropy error is decreased at the cost of increasing computation time in this Monte Carlo 1D measurement ray expected entropy case.}
\label{fig:ApproxJust}
\end{figure}

\section{Future Pose Optimization}

In this section, we show how expected entropy changes from single measurement rays can be integrated into predicting the expected information gain of measurement scans. This metric, and a cost associated with travel time, are integrated to determine an optimal future pose and collision-free path in a 2D environment.

\subsection{Optimal Attitude from Sample Rays}

The objective is to choose a future pose that minimizes the map uncertainty, such that the robot can autonomously move toward this optimal goal. The robot pose is selected among various positions and attitudes. Since searching over all locations and attitudes would require infinite computational resources, these search spaces are discretized into a limited number of robot attitudes and positions.
The procedure involves three steps. First, the attitude that maximizes the information gain objective function is selected at each candidate future pose location. Second, the location with the maximum objective function of all candidates is selected. Thus, the optimal attitude and the optimal location compose the optimal future pose. Finally, the robot determines and follows a collision-free path to the optimal future pose. This process is repeated, resulting in autonomous exploration.

\paragraph{Attitude Optimization}
Consider a collision-free pose candidate at an arbitrary location $x_c$. At this pose, consider $n_d$ evenly-spaced measurement rays oriented radially outward from the robot in a circular pattern (see red lines in Figure \ref{fig:OptProcess}). These rays are denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$. Attitudes correspond to the rays such that the robot sensor is aligned with the ray direction, and these are denoted $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor field of view (FOV). We choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
\begin{align}
\label{eqn:FindRc}
&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad\quad\quad\quad\quad
-\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,i}))]\bigg).
%\mathcal I_\text{ray}(x_c,z_{c,i})\right).
\end{align}

This method provides the attitude that maximizes the information gain at an arbitrary location in 2D space. The set of candidate positions that warrant consideration is determined with one of two methods, namely \emph{expanding ring} and \emph{complete Cartesian}, described next.

\paragraph{Expanding Ring}
The expanding ring technique is advantageous for searching local solutions quickly, only considering distant future poses when necessary. The key idea is that future candidate locations lie on a circular ``ring'' centered around the robot, evenly spaced around the ring (see red circles in Figure \ref{fig:OptProcess}), and the ring is expanded if certain criteria are not met. More explicitly, consider $n_c$ candidate locations denoted by $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$ located with the distance $\delta$ away from the robot. Any location that violates inequality constraint \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions. The set of optimal attitudes at each candidate location is $\braces{R^*_{1},R^*_{2},\ldots,R^*_{n_c}}$, obtained from \refeqn{FindRc}. The objective function \refeqn{ObjFun} is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
\begin{align}
\label{eqn:ObjFunApprox}
\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*,z_{c,i}))\right]\bigg),
\\
x_c^*&=\argmax_{x_c}\mathcal I(x_c,R_c^*).
\end{align}
At the optimal pose $X^*_c=(x^*_c,R^*_c)$, the resulting information gain must satisfy $\mathcal I(X_c^*)\geq\mathcal I_\text{min}$, where $\mathcal I_\text{min}$ is some minimum threshold for expected information gain; robot motion is only justified when the expected information gain is significantly large. If the minimum threshold is not met, the ring of candidate pose locations is increased such that $n_c$ and $\delta$ are multiplied by a scale function $\lambda>1$.
This process is repeated until the expected information gain of the optimal pose is at least $\mathcal I_\text{min}$, or all candidates lie in collision zones or outside map limits. %A pseudocode of the process is found in Table \ref{tab:Alg_AutomousExploration}.

The primary advantage of the expanding ring approach to search the 2D space is that the number of pose candidates need not be proportional to the map area, and that local maxima tend to fall within short distances of the current robot pose. The main drawback is that poses outside a small local region of the robot are frequently neglected, and the robot frequently executes repeated short motions, often without large information gains.

\paragraph{Complete Cartesian}
The complete Cartesian method to search the 2D space provides candidates located fixed distance $d$ apart in each Cartesian direction. This approach is advantageous for capturing expected information gains in regions outside of the immediate vicinity of the robot, but must be applied carefully to avoid issues with computational bottlenecks due to the potentially-large search space.

Similar to the expanding ring approach, $n_c$ candidate pose locations are considered, where those violating \refeqn{CollisionInequalityConstraint} are neglected from further consideration. However, unlike the ring expanding approach, the candidates have different distances from the robot in general, so the objective function is modified to enforce a cost on the squared distance from the current pose location $x_t$ to the candidate location $x_c$, i.e.,
\begin{align}
\label{eqn:ObjFunApproxCompleteCartesian}
\mathcal I&(x_c,R_c^*,x_t)\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*(x_c),z_{c,i}))\right]\bigg)-k_\text{dist}(x_t-x_c)^\T(x_t-x_c),
\end{align}
where weighting function $k_\text{dist}$ represents the sensitivity to avoiding large motions across the map. Furthermore, the same candidate locations are considered throughout the repeated processes of exploration, and the map cell occupancies are assumed static. Therefore, if candidate location $x_c$ fails to satisfy $\mathcal I(x_c,R_c^*,x_t)\geq\mathcal I_\text{min}$, then $x_c$ need \emph{never} be considered again. Avoiding these unnecessary calculations yields a greatly lowered computational burden, particularly late in exploration where several regions are collision-free but disadvantageous to revisit.

\begin{figure}
\vspace*{0.1\columnwidth}
\centerline{
	\includegraphics[height=0.35\columnwidth]{ExampleOptimalPose.png}
}
\begin{picture}(0,0)(0,0)
\setlength{\unitlength}{0.1\columnwidth}\scriptsize
\put(4.4,2.1){\color{blue}$X_t$}
\put(5.9,1.9){\color{red}$x_1$}
\put(2.9,2.0){\color{red}$x_5$}
\put(4.4,3.6){\color{red}$x_3$}
\put(4.4,0.6){\color{red}$x_7$}
\put(5.4,3.0){\color{red}$x_2$}
\put(3.3,3.1){\color{red}$x_4$}
\put(3.3,1.0){\color{red}$x_6$}
\put(5.4,1.0){\color{red}$x_8$}
\put(6.8,3.1){\color{red}$z_{2,1}$}
\put(4.5,3.1){\color{red}$z_{2,5}$}
\put(5.6,4.1){\color{red}$z_{2,3}$}
\put(5.6,2.2){\color{red}$z_{2,7}$}
\put(6.5,3.7){\color{red}$z_{2,2}$}
\put(4.9,3.8){\color{red}$z_{2,4}$}
\put(5.0,2.5){\color{red}$z_{2,6}$}
\put(6.5,2.5){\color{red}$z_{2,8}$}
\put(6.1,2.9){$X_c^*$}
\end{picture}
\caption{This figure illustrates the expanding ring method. Initially, a robot (blue circle) views down and left (blue sector). Then, the robot considers $n_c=8$ poses (red circles) as possible candidates for where to move next. For each candidate, $n_d=8$ directions are considered (red lines, only displayed on candidate location $c=2$). Then, the expected entropy of each ray is calculated. The scan (red sector) covering those rays with the largest expected entropy decreases is chosen for each candidate, and the best candidate $X_c^*$ (black circle: location, black sector: scan) is chosen to maximize information gain. Finally, Dijkstra's algorithm provides the collision-free motion between $X_t$ and $X_c^*$, and the process is repeated.
}
\label{fig:OptProcess}
\end{figure}

%Then, \refeqn{DiscExpEntropyRay} is applied repeatedly over several sample measurement rays that compose scan $Z_c$ as shown in~\cite{KauAiLee16}.

\subsection{Travel Cost with Obstacles}
% Dijkstra's cost map

\subsection{Optimal Pose Selection}
% Single-Vehicle argmax

\subsection{Collision-Free Motion Planning}
% Dijkstra's algorithm (after cost map), no least squares yet

\section{Numerical Examples}

\subsection{Exploring a Simple Environment}
% SIMPAR numerical example (2 rooms, 1 hallway)

\subsection{Exploring a Complicated Benchmark Environment}
% JINT17 Fig. 7 & 8

\section{Experimental Result}
% JINT17 9, 10, 11 (ground experiment at the NRL)

\section{Conclusions}