### 2018-07-01 Dissertation/Thesis/tex/conclusions.tex
added beginning of 3D exploration results, made template for conclusions

\chapter{Conclusions}\label{chap:Conclusions}

This dissertation proposed several contributions in probabilistic occupancy grid mapping and autonomous exploration in 2D and 3D environments.