### 2018-07-13 Dissertation/Defense/defense.tex
Acknowleged the NRL

\documentclass[11pt,professionalfonts,hyperref={pdftex,pdfpagemode=none,pdfstartview=FitH}]{beamer}
%\usepackage{times}
%\usefonttheme{serif}
%\usepackage{helvet}
%\usepackage{amsmath,amssymb}
\usepackage{graphicx,multirow}

%\usepackage{movie15}
\usepackage{media9}

\usepackage{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}

%\usepackage{warmread}
%\usepackage[all,import]{xy}

%\renewcommand\mathfamilydefault{\rmdefault}

% JIRS includes
%\usepackage{graphicx}
%\usepackage{amsmath,amssymb,url,times}%,subfigure}% amsthm is the one!
%\usepackage{caption,subcaption,hyperref}
%\usepackage{color,comment}
%\usepackage{curves,pgfgantt}

\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}
\newcommand{\bracket}[1]{\ensuremath{\left[ #1 \right]}}
\newcommand{\braces}[1]{\ensuremath{\left\{ #1 \right\}}}
\newcommand{\parenth}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\pair}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\met}[1]{\ensuremath{\langle\langle #1 \rangle\rangle}}
\newcommand{\refeqn}[1]{(\ref{eqn:#1})}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\tr}[1]{\mathrm{tr}\ensuremath{\negthickspace\bracket{#1}}}
\newcommand{\trs}[1]{\mathrm{tr}\ensuremath{[#1]}}
\newcommand{\deriv}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\newcommand{\SO}{\ensuremath{\mathsf{SO(3)}}}
\newcommand{\T}{\ensuremath{\mathsf{T}}}
\renewcommand{\L}{\ensuremath{\mathsf{L}}}
\newcommand{\so}{\ensuremath{\mathfrak{so}(3)}}
\newcommand{\SE}{\ensuremath{\mathsf{SE(3)}}}
\newcommand{\se}{\ensuremath{\mathfrak{se}(3)}}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\aSE}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&1\end{bmatrix}}}
\newcommand{\ase}[2]{\ensuremath{\begin{bmatrix}#1&#2\\0&0\end{bmatrix}}}
\newcommand{\D}{\ensuremath{\mathbf{D}}}
\newcommand{\Sph}{\ensuremath{\mathsf{S}}}
\renewcommand{\S}{\Sph}
\newcommand{\J}{\ensuremath{\mathbf{J}}}
\newcommand{\Ad}{\ensuremath{\mathrm{Ad}}}
\newcommand{\intp}{\ensuremath{\mathbf{i}}}
\newcommand{\extd}{\ensuremath{\mathbf{d}}}
\newcommand{\hor}{\ensuremath{\mathrm{hor}}}
\newcommand{\ver}{\ensuremath{\mathrm{ver}}}
\newcommand{\dyn}{\ensuremath{\mathrm{dyn}}}
\newcommand{\geo}{\ensuremath{\mathrm{geo}}}
\newcommand{\Q}{\ensuremath{\mathsf{Q}}}
\newcommand{\G}{\ensuremath{\mathsf{G}}}
\newcommand{\g}{\ensuremath{\mathfrak{g}}}
\newcommand{\Hess}{\ensuremath{\mathrm{Hess}}}
\newcommand{\refprop}[1]{Proposition \ref{prop:#1}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

\definecolor{mygray}{gray}{0.9}

\graphicspath{{../../Fig/}}

\mode<presentation> {
  \usetheme{Warsaw}
  \usefonttheme{serif}
  \setbeamercovered{transparent}
}

\newcommand{\mypaper}{}

\setbeamertemplate{footline}%{split theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.7\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fill]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm]{title in head/foot}
    \usebeamerfont{title in head/foot}\mypaper\hfill \insertframenumber/\inserttotalframenumber
    \usebeamerfont{title in head/foot}\hfill \insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
} \setbeamercolor{box}{fg=black,bg=yellow}

\title[Multi-Robot Probabilistic Mapping and Exploration]{\large Multi-Robot Probabilistic Mapping and Exploration}

\author{\vspace*{-0.1cm}}

\institute{\footnotesize
{\normalsize Evan Kaufman}\\
\vspace*{0.3cm}
Mechanical and Aerospace Engineering\\George Washington University
}

%\institute{\footnotesize
%{\normalsize Evan Kaufman\\Research Advisor: Taeyoung Lee}\vspace*{0.1cm}\\
%  Mechanical and Aerospace Engineering\\ George Washington University \vspace*{0.3cm}\\ {\normalsize Special Thanks to:\\
% \vspace*{0.2cm}
% Kuya Takami, Mahdis Bisheban, and Kanishke Gamagedara} \\Department of Mechanical Engineering\\George Washington University\\
%\vspace*{0.2cm}
%{\normalsize Zhuming Ai, Ira. S. Moskowitz, and Mark Livingston}\\Information Management \& Decision Architectures\\U.S. Naval Research Laboratory}

\date{}

\definecolor{tmp}{rgb}{0.804,0.941,1.0}
\setbeamercolor{numerical}{fg=black,bg=tmp}
\setbeamercolor{exact}{fg=black,bg=red}

\newtheorem{prop}{Proposition}

\renewcommand{\emph}[1]{\textit{\textbf{\color{blue}{#1}}}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{}
\subsection*{Motivation}

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
	\item Mobile Robotics
	\begin{itemize}
		\item Environments can be dangerous and uncertain
		\item Robots can replace humans for tasks such as surveillance, search-and-rescue, and cleaning
	\end{itemize}
	\pause
	\item Robotic Capabilities
	\begin{itemize}
		\item Robot must understand their environment by generating a map, referred to as \emph{mapping}
		\item Selecting and performing robotic actions to maximize map information, referred to as \emph{autonomous exploration}, is necessary to optimally explore initially-uncertain environments
		\item Cooperative robots can share information to achieve a collective goal with \emph{multi-vehicle exploration and patrol}
		\item These tasks must be performed in 2D and 3D environments, where computational efficiency is paramount to real-time implementation
	\end{itemize}
\end{itemize}
\end{frame}

\subsection*{Research Contributions}

\begin{frame}
\frametitle{Research Contributions}
\begin{itemize}
	\item Occupancy Grid Mapping
		\begin{itemize}
		\item Computation of the exact occupancy probability
		\item Incorporation of prior knowledge and new measurements
		\item Bayesian information fusion
	\end{itemize}
	\pause
	\item Autonomous Exploration
	\begin{itemize}
		\item Novel approach to predicting future map uncertainty
		\item Optimal planning to maximize information gain
	\end{itemize}
	\pause
	\item Multi-Vehicle Exploration and Patrol
	\begin{itemize}
		\item Bidding-based framework for assigning tasks while exploring uncertain environments
		\item Map degradation to encourage revisiting regions from the past 
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Outline}
\begin{enumerate}[I]
	\item Occupancy Grid Mapping
	\begin{itemize}
		\item Background and problem definition
		\item Common approaches and approximations
		\item Inverse sensor model in 1D
		\item Ray casting and measurement scans for 2D and 3D environments
		\item Numerical examples
	\end{itemize}
	\pause
	\item Autonomous Exploration
	\begin{itemize}
		\item Existing solutions: frontiers, expected measurements
		\item Proposed approach: entropy expectation
		\item Collision-free trajectories
		\item Optimal pose selection
		\item Numerical examples
	\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Outline}
\begin{enumerate}[I]
\addtocounter{enumi}{2}
	\item Multi-Vehicle Exploration and Patrol
	\begin{itemize}
		\item Cooperative approaches
		\item Bidding-based framework
		\item Receding horizon
		\item Multi-vehicle autonomous exploration numerical example
		\item Autonomous patrol
		\item Multi-vehicle autonomous patrol numerical example
	\end{itemize}
	\pause
	\item Experiments
	\begin{itemize}
		\item Ground vehicle 2D mapping and autonomous exploration
		\item Aerial quadrotor in large vertically-uniform space
		\item Aerial quadrotor in complex 3D environment
	\end{itemize}
\end{enumerate}
\end{frame}

\section*{}
\subsection*{Probabilistic Occupancy Grid Mapping}

\begin{frame}
\frametitle{Probabilistic Occupancy Grid Mapping}
\begin{enumerate}[I]
	\item Probabilistic Occupancy Grid Mapping
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Mapping Representation}
    \begin{itemize}
    	\item Simple and well-known grid-based representation
	\pause
	\item Decomposes space into evenly-spaced cells that are either \emph{occupied} or \emph{free}
	\pause
    	\item Probabilistic: goal is to find the \emph{probability} that the cells are occupied
	\pause
	\item Useful for collision-avoidance and graph searching
    \end{itemize}

    \only<1->{
\begin{figure}
\centerline{
    \includegraphics[height=2.1cm]{ogm_ex3.jpeg}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex1.jpg}\hspace*{0.1cm}
\hspace*{0.5cm}
    \includegraphics[height=2.1cm]{ogm_ex2.png}\hspace*{0.1cm}
}
\end{figure}}

\end{frame}

\begin{frame}
\frametitle{Problem Definition}
%\framesubtitle{Problem Definition}

\begin{itemize}
	\item The Map and the Robot
	\begin{itemize}
	\item Map $m$ is composed of $n_m$ grid cells with known location and size
	\item The $i$-th grid cell $\mathbf{m}_i$ is a \emph{static binary} random variable, independent of other grid cells: $P(m)=P(\mathbf{m}_1,\mathbf{m}_2,\ldots,\mathbf{m}_{n_m})=\prod_{i=1}^{n_m}P(\mathbf{m}_i)$
	\item \emph{Pose} $X_t$ is known, containing robot \emph{position} and \emph{attitude}
	\end{itemize}
\vspace*{0.0cm}\pause
\end{itemize}
\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Depth Measurements
	\begin{itemize}
	\item Each measurement origin and direction is known \emph{deterministically}
	\item A measurement \emph{scan} $Z_t=\braces{z_{t,1},z_{t,2},\ldots,z_{t,n_z}}$ contains $n_z$ measurement \emph{rays} (depths)% at the $t$-th time step, and the history of measurement scans $Z_{1:t}$ is known
\item The \emph{forward sensor model} is known from the sensor properties
\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
%The \emph{forward sensor model} is the probability density distribution $p(z_{t,l}|m,X_{t})$  known from the sensor properties
\hspace*{0.25cm}
\begin{figure}[!htbp]
\vspace*{-0.25cm}
%\vspace*{0.25cm}
\centerline{
    \includegraphics[width=3.5cm]{BeamModel.png}\hspace*{0.1cm}
    }
\vspace*{0.25cm}
\centerline{
    \includegraphics[width=2.5cm]{1D_True_Grid.png}\hspace*{0.1cm}
%    \hspace*{0.75cm}
    }
{Beam Model for Range Finders}
\end{figure}

\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Problem Definition}

\begin{itemize}
	\item Bayesian Framework
	\begin{itemize}
	\item Markov Assumption: latest a priori cell occupancy probabilities capture the information from all prior observations
	\item Log-Odds Ratio: popular representation due to simple additive update structure and probability truncation avoidance, but requires the \emph{assumption}
	\begin{align*}
		P(Z_t|\mathbf{m}_i,X_{1:t},Z_{1:t-1})\approx P(Z_t|\mathbf{m}_i,X_t)
	\end{align*}
	\end{itemize}
\vspace*{0.0cm}\pause
	\item Inverse Sensor Model
	\begin{align*}
&P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\eta_{t,l}\sum_{m\in\mathcal{M}_i}p(z_{t,l}|m,X_{t})P(m|X_{1:t-1},Z_{1:t-1}).
\end{align*}
	\begin{itemize}
	\item Given $n$ grid cells: $\mathcal O(2^n)$ is \emph{computationally-intractable}, motivating a different solution
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Approximate Solutions}

\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Approximate a function for the inverse sensor model based on intuition
	\begin{itemize}
		\item Simple to implement, but mathematically inaccurate
	\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
\begin{figure}[!htbp]
\centerline{
    \vspace*{-0.5cm}
%	\hspace*{0.25cm}
    \includegraphics[width=4.0cm]{Approx_ISM_shortened.png}\hspace*{0.1cm}
    }
\centerline{
    \includegraphics[width=3.5cm]{1D_True_Grid.png}
%    \hspace*{0.75cm}
    }
		\end{figure}
\end{minipage}
\vspace*{-0.5cm}
\vspace*{0.0cm}\pause
\begin{itemize}
	\item Find a solution through learning
	\begin{itemize}
		\item Simulate maps, poses, and measurements and use learning to obtain an inverse sensor model
		\item Complicated, unclear how these parameters are chosen
	\end{itemize}
	\vspace*{0.0cm}\pause
	\item Goal: design a \emph{simple} and \emph{accurate} occupancy grid mapping method avoiding log-odds ratio assumptions
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Proposed Approach: Utilize Invariance}

\begin{itemize}
    \item Main idea: make use of occupancy grid mapping \emph{assumptions} and extract \emph{patterns} from probabilistic properties to find a \emph{computationally-efficient} solution
	\begin{itemize}
		\item Since the origin and direction of each measurement ray is known deterministically, the set of grid cells that the ray intersects is \emph{known through geometry}
		\item A depth reading follows the forward sensor model, which \emph{only depends} on the first occupied grid cell along the measurement ray
	\end{itemize}
\end{itemize}
\setcounter{subfigure}{0}
% uncomment
\begin{figure}
  \centering
  \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_1.png}
  \end{subfigure}
  \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_2.png}
  \end{subfigure}
    \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_3.png}
  \end{subfigure}
  \begin{subfigure}[t]{.45\linewidth}
    \centering\includegraphics[width=\linewidth]{rkplus_4.png}
  \end{subfigure}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Bayesian Update}

\begin{itemize}
	\item Unnormalized Probability:
	\begin{align*}
\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})
&=\mathbf{P}_k^-
\bigg[\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]\nonumber\\
&\qquad + \bigg\{\prod_{j=0}^{k-1}\bar{\mathbf{P}}_j^-\bigg\}p(z|\mathbf{r}_{k+},X_t)\mathbf{P}_k^-
\end{align*}
	\item Normalizer:
	\begin{align*}
\label{eqn:allEta}
\eta
&=
\bigg[\sum_{i=1}^{n_{r}+1}\bigg\{\prod_{j=0}^{i-1}\bar{\mathbf{P}}_j^-\bigg\} p(z|\mathbf{r}_{i+},X_t)\mathbf{P}_k^-\bigg]^{-1}
\end{align*}
	\item Solution: $P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})=\eta\tilde P(\mathbf{r}_{k}|z,X_{1:t},Z_{1:t-1})$
	\item \emph{Linear Complexity}: $\mathcal{O}(n)$ for \emph{all} $n$ cells in FOV compared with $\mathcal O(n\times2^n)$ without using invariance
\end{itemize}
\end{frame}

% Ray Casting

\begin{frame}
\frametitle{Ray Casting}

\begin{itemize}
	\item Required for 2D and 3D occupancy grid maps
	\item Main idea: use simple geometry to find the distance from the robot sensor to each cell intersected by the measurement ray
\end{itemize}

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[t]{0.4\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{RayCastingIllustration.png}
%        \label{fig:penn_map_total}
    \end{subfigure}
    \hspace*{0.05\columnwidth}
    \begin{subfigure}[t]{0.4\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{RayCastIllustrationReducedMapOnly.png}
%        \label{fig:penn_map_zoom}
    \end{subfigure}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Scan Combination Approaches}

\begin{itemize}
	\item \emph{Ray-By-Ray Update}: given a scan of measurement rays, update the map based on each measurement ray \emph{individually} and \emph{sequentially}
	\begin{itemize}
		\item Benefit: the measurement rays maintain some dependency as they capture the same map
		\item Drawback: the best order for evaluating rays is not necessarily chosen
	\end{itemize}
%	\item Conditional probability:
%	\begin{align*}
%		P(\mathbf{m}_i|X_{1:t}&,Z_{1:t})\nonumber\\&
%		=P((\dots(((\mathbf{m}_i|X_{1:t},Z_{1:t-1})|z_{t,1})|z_{t,2})|\ldots)|z_{t,n_z})
%	\end{align*}
	\item \emph{Synergistic Update}: consider every measurement ray inside a scan \emph{simultaneously} to update all grid cells inside the FOV
	\begin{itemize}
		\item Benefit: synergistic update provides a method where all rays are considered simultaneously
		\item Drawback: required assumption that the measurement rays are independent, though they capture the same map
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Numerical Example}

\begin{itemize}
	\item A standard \emph{approximate} inverse sensor model technique is compared with the proposed \emph{exact} method
	\item The exact algorithm yields a more \emph{certain} map from the same measurement set
\end{itemize}

%\vspace*{0.5cm}
\begin{figure}[!htbp]
    \centering
    \begin{subfigure}{0.2\textwidth}
        \centering
        \includegraphics[width=\textwidth]{AISM_Image_19.pdf}
        \caption*{Approx. Map}
    \end{subfigure}
%    \hspace*{-0.1\textwidth}
    \begin{subfigure}{0.2\textwidth}
        \centering
        \includegraphics[width=\textwidth]{EISM_Image_19.pdf}
        \caption*{Exact Map}
    \end{subfigure}
    \begin{subfigure}{0.2\textwidth}
        \centering
        \includegraphics[width=\textwidth]{AISM_Image_inf_19.pdf}
        \caption*{Approx. $H$}
    \end{subfigure}
%    \hspace*{-0.1\textwidth}
    \begin{subfigure}{0.2\textwidth}
        \centering
        \includegraphics[width=\textwidth]{EISM_Image_inf_19.pdf}
        \caption*{Exact $H$}
    \end{subfigure}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Probabilistic Occupancy Grid Mapping Summary}

\begin{itemize}
	\item The proposed novel Bayesian solution to the inverse sensor model provides the exact solution without approximate functions or learned solutions
	\item The computational complexity is linear with respect to the number of cells considered, making real-time implementation possible
	\item 2D and 3D maps are updated with ray casting and ray-by-ray combination
\end{itemize}

\end{frame}

\section*{}
\subsection*{Entropy-Based Autonomous Exploration}

\begin{frame}
\frametitle{Entropy-Based Autonomous Exploration}
\begin{enumerate}[I]
\addtocounter{enumi}{1}
	\item Entropy-Based Autonomous Exploration
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Autonomous Exploration}

\begin{itemize}
	\item Goal: determine robotic actions that maximize map information
	\item Approach: predict future map uncertainty and travel distance to select an optimal pose
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Existing Exploration Approaches}

\begin{itemize}
	\item Frontier-Based Exploration
	\begin{itemize}
		\item A robot identifies boundaries of free and uncertain cells
		\item The robot moves toward these frontiers, thereby pushing back the boundaries
		\item Heuristic, suboptimal
	\end{itemize}
	\vspace*{0.0cm}\pause
	\item Entropy-Based Approaches
	\begin{itemize}
		\item Probabilities from the occupancy grid map are approximate
		\item Usage of ``hallucination measurements'': assume that $\text{E}[H(P(\mathbf{m}_i|z))]\approx H(P(\mathbf{m}_i|\text{E}[z]))$
	\end{itemize}
	\vspace*{0.0cm}\pause
	\item Proposed Approach
	\begin{itemize}
		\item Use occupancy grid map probabilities
		\item Solve $\text{E}[H(P)]$ and select future poses to maximize map information
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Entropy Definition}
\begin{itemize}
        	\item Uncertainty-Based Exploration
	\begin{itemize}
		\item Main idea: choose robot motions that minimize the uncertainty of the occupancy grid map
		\item Equivalently maximize map information gain
	\end{itemize}
\end{itemize}
\begin{minipage}[t]{7.0cm}
\begin{itemize}
	\item Entropy
	\begin{itemize}		
		\item \emph{Shannon's entropy} serves as an uncertainty measure
		\begin{align*}
			H(P)=-P\log P-\bar{P}\log \bar{P}%-(1-P)\log(1-P)
		\end{align*}
		with cell occupancy probability $P$ and complement $\bar{P}=1-P$
		\item Entropy $H$ is maximized when $P=0.5$ and minimized as $P\rightarrow0$ or $P\rightarrow1$
	\end{itemize}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{3.0cm}
\vspace*{0.5cm}
\begin{figure}[!htbp]
	\centerline{
		\hspace*{1.25cm}
   		\includegraphics[width=5.0cm]{H_Plotted_square.pdf}%\hspace*{0.1cm}
	}
\end{figure}
\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Expected Information Gain}

\begin{itemize}
	\item Main Idea: determine the benefit of future poses based on how their measurements are expected to improve the map
	\vspace*{0.0cm}\pause
	\item Goal: decrease total map entropy:
	\begin{align*}
		H(P(m))&=\sum_{i=1}^{n_m}H(P(\mathbf{m}_i))
	\end{align*}
	\vspace*{0.0cm}\pause
	\item Measurement Ray Expected Information Gain:
	\begin{align*}
		\text{E}[H(P(m|x_c,z_{c}))]&=\sum_{k=1}^{n_{r}+1}\bigg\{H(P(m|x_c,z_{c,k}))P(z_{c,k}|x_c)\bigg\}
		\\
		P(z_{c,k}|x_c)&=\frac{p(z_{c,k}|x_c)}{\sum_{i=1}^{n_{r}+1}p(z_{c,i}|x_c)}=\frac{\eta_{c,k}^{-1}}{\sum_{i=1}^{n_{r}+1}\eta_{c,i}^{-1}}
		\\
		\mathcal I(x_c,z_{c})&=H(P(m))-\text{E}[H(P(m|x_c,z_{c}))]
	\end{align*}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Expected Information Gain Computation}

\begin{itemize}
	\item \emph{Squared Complexity}: $\mathcal O(n_r^2)$ for $n_r$ cells along the ray
	\vspace*{0.0cm}\pause
	\item Two Approaches:
	\begin{itemize}
		\vspace*{0.0cm}\pause
		\item Decrease the cells considered by selecting $n_p$ most probable cells: $\mathcal O(n_r^2)$ reduces to $\mathcal O(n_p^2)$
		\vspace*{0.0cm}\pause
		\item Use 3D cells in a 2D \emph{projection}
\begin{figure}
  \centering
  \begin{subfigure}[t]{.3\linewidth}
    \centering\includegraphics[height=.75\linewidth]{experiment_ogm3D_2min47sec.jpg}
    \caption*{\ \ \ \ \ \ \ \ \ \ \ \ \ \ 3D Map}
  \end{subfigure}
  \hspace*{0.25\linewidth}
  \begin{subfigure}[t]{.3\linewidth}
    \centering\includegraphics[height=.75\linewidth]{experiment_2min47sec.jpg}
    \caption*{2D Projection}
  \end{subfigure}
\end{figure}
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Pose Selection}

\begin{itemize}
	\item Main Idea: select the future pose that maximizes map information gain while accounting for travel distance
	\vspace*{0.0cm}\pause
	\item Attitude Selection: choose the attitude that covers the candidate measurement rays with largest objective function summation
\begin{figure}
\centerline{
	\includegraphics[height=0.35\linewidth]{ExampleOptimalPose.png}
%\begin{picture}(0,0)(0,0)
%\setlength{\unitlength}{0.1\linewidth}\scriptsize
%\put(4.4,2.1){\color{blue}$X_t$}
%\put(5.9,1.9){\color{red}$x_1$}
%\put(2.9,2.0){\color{red}$x_5$}
%\put(4.4,3.6){\color{red}$x_3$}
%\put(4.4,0.6){\color{red}$x_7$}
%\put(5.4,3.0){\color{red}$x_2$}
%\put(3.3,3.1){\color{red}$x_4$}
%\put(3.3,1.0){\color{red}$x_6$}
%\put(5.4,1.0){\color{red}$x_8$}
%\put(6.8,3.1){\color{red}$z_{2,1}$}
%\put(4.5,3.1){\color{red}$z_{2,5}$}
%\put(5.6,4.1){\color{red}$z_{2,3}$}
%\put(5.6,2.2){\color{red}$z_{2,7}$}
%\put(6.5,3.7){\color{red}$z_{2,2}$}
%\put(4.9,3.8){\color{red}$z_{2,4}$}
%\put(5.0,2.5){\color{red}$z_{2,6}$}
%\put(6.5,2.5){\color{red}$z_{2,8}$}
%\put(6.1,2.9){$X_c^*$}
%\end{picture}
}
\vspace*{-0.05\linewidth}
\end{figure}
	\item The attitude $R_c$ is selected for all candidates: the expected information gain becomes $\mathcal I(X_c)$ where $X_c=\braces{x_c,R_c}$
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Pose Selection}

\begin{itemize}
	\item Travel Distance: generate a cost map from the occupancy grid to obtain distances to each candidate pose $d(x_c)$
	\vspace*{0.0cm}\pause
	\item Bump Function: use a bump function \begin{align*}f(d)=\exp\braces{-\frac{d^2}{2\sigma^2}}+\gamma, \quad \sigma>0, \quad \gamma>0\end{align*} to incentivize the robot to explore more locally until nearby spaces are well-known
	\begin{figure}
	\vspace*{-0.015\linewidth}
\centering
	\includegraphics[width=0.8\linewidth]{GaussianBumpFunFlat.png}
		\vspace*{-0.015\linewidth}
\end{figure}
	% C++: exp(-0.5*distAway*distAway/sigSqrd)+minFunVal GaussianBumpFunFlat.png
	\vspace*{0.0cm}\pause
	\item Select the pose that maximizes the product \begin{align*}X^*=\argmax_{X_c}{\mathcal I(X_c)f(d(x_c))}\end{align*}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Collision-Free Path}

\begin{itemize}
	\item Using the cost map from before, \emph{Dijkstra's algorithm} is completed by finding the waypoints from the desired candidate back to the current location
	\vspace*{0.0cm}\pause
	\item These waypoints serve as input to a \emph{constrained polynomial least squares trajectory}
	\begin{itemize}
		\item Each trajectory is solved independently as a function of time, i.e. $x(t)$, $y(t)$
		\item Starting and terminal locations are fixed
		\item Segments patched together have the same position and velocity at the connection
	\end{itemize}
	\vspace*{0.0cm}\pause
	\item A \emph{geometric controller} tracks the trajectory without \emph{singularities} or \emph{ambiguities}
	\vspace*{0.0cm}\pause
	\item The mapping, exploration, and control run in \emph{real-time}
\end{itemize}

\end{frame}

\begin{frame}
	\frametitle{2D Autonomous Exploration: Intel Research Lab}
	\centering{
		\includemedia[
  		width=0.720\textwidth,
  		height=0.576\textwidth,
  		activate=pageopen,
  		addresource=IRL.mp4,
  		flashvars={source=IRL.mp4}
		]{}{VPlayer.swf}
		}
\end{frame}

\begin{frame}
	\frametitle{3D Autonomous Exploration: Mars}
	\centering{
		\includemedia[
  		width=0.720\textwidth,
  		height=0.576\textwidth,
  		activate=pageopen,
  		addresource=Topological Mars Mapping and Autonomous Exploration in 3D.mp4,
  		flashvars={source=Topological Mars Mapping and Autonomous Exploration in 3D.mp4}
		]{}{VPlayer.swf}
		}
\end{frame}

\begin{frame}
\frametitle{Autonomous Exploration Summary}

\begin{itemize}
	\item Other approaches use heuristic frontiers or entropy of expected measurements
	\item The proposed approach solves for the expected value of future map entropy directly
	\item Several sample future measurement rays and Dijkstra's cost map combine in an objective function for optimality
	\item Numerical simulations show a large 2D environment, and this approach can be extended to 3D environments
\end{itemize}

\end{frame}

\section*{}
\subsection*{Multi-Vehicle Exploration and Patrol}

\begin{frame}
\frametitle{Multi-Vehicle Exploration and Patrol}
\begin{enumerate}[I]
\addtocounter{enumi}{2}
	\item Multi-Vehicle Exploration and Patrol
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Motivation and Goal}

\begin{itemize}
	\item Multi-Vehicle Autonomous Exploration
	\begin{itemize}
		\item Multiple agents can accomplish tasks faster and more reliably than a single agent
		\item Large groups are capable of coverage in large environments
		\item Problem: multi-vehicle map information gain optimality can be computationally-intractable
		\item Approach: bidding-based solution simplifies a complicated optimization into a series of simplified optimizations
	\end{itemize}
	\pause
	\item Multi-Vehicle Autonomous Patrol
	\begin{itemize}
		\item Patrolling large environments with humans is expensive and sometimes dangerous
		\item Several robotic approaches introduce new variables for patrol
		\item Approach: degrade probabilistic occupancy grid over time to encourage revisiting past regions
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Related Work in Multi-Vehicle Exploration}

\begin{itemize}
	\item Most existing bidding-based techniques deal with auctioning tasks among a few agents locally in decentralized frameworks
	\item The other approach is for a central executive to run a series of auctions
	\item Bids are discounted between auctions to coordinate the efforts based on coverage overlap
	\item Heuristic, ad hoc
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Proposed Approach in Multi-Vehicle Exploration}

\begin{itemize}
\item Central executive completes $n$ auctions for $n$ robots
\begin{enumerate}
	\item Each robot determines its optimal future pose, and bids the objective function
	\item The winner is awarded this future pose, and the remaining robots update their bids according to
	\begin{itemize}
		\item New expected map changes from the winner
		\item Collision-avoidance with the winner
	\end{itemize}
	\item Repeat these steps with the remaining robots
\end{enumerate}
\item The efforts are coordinated by updating bids between auctions
\item The algorithm follows a receding-horizon framework: future pose optimization repeats as quickly as possible
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Numerical Example: SEH Multi-Vehicle Exploration}

	\centering{
		\includemedia[
  		width=0.720\textwidth,
  		height=0.576\textwidth,
  		activate=pageopen,
  		addresource=Multi-Vehicle Mapping and Exploration.mp4,
  		flashvars={source=Multi-Vehicle Mapping and Exploration.mp4}
		]{}{VPlayer.swf}
		}

\end{frame}

\begin{frame}
\frametitle{Cooperative Autonomous Patrol}

\begin{itemize}
	\item Existing Approaches
	\begin{itemize}
		\item Rely on a known map and use Voronoi diagrams to disperse
		\item Another approach introduces cell periodicity: observe long-term dynamics of objects at the same location on the occupancy grid
		\item Bio-inspired approach uses digital pheromones 
		\item Minimizing robot idleness
	\end{itemize}
	\pause
	\item Proposed Approach
	\begin{itemize}
		\item Degrade occupancy grid cells toward their initial values with exponential decay
		\item Only affects existing probability values without new variables to track
		\item Promotes exploration first, then patrol
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Continuous-Time Markov Process}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{markov_diagram_continuous.pdf}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{DegradeExamples.pdf}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Numerical Example: SEH Multi-Vehicle Patrol}

	\centering{
		\includemedia[
  		width=0.720\textwidth,
  		height=0.576\textwidth,
  		activate=pageopen,
  		addresource=Multi-Vehicle Cooperative Patrol.mp4,
  		flashvars={source=Multi-Vehicle Cooperative Patrol.mp4}
		]{}{VPlayer.swf}
		}

\end{frame}

\begin{frame}
\frametitle{Cell Degradation and Patrol}

\begin{figure}
	\centering
	\includegraphics[width=0.65\textwidth]{entropy_patrol_9minSwitch.pdf}
\end{figure}

\begin{itemize}
	\item Entropy of the map increases if patrol stops at $9$ minutes if cell degradation continues
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Multi-Vehicle Exploration and Patrol Summary}

\begin{itemize}
	\item Robotic tasks are assigned following a bidding-based framework to avoid a computationally-intractable solution
	\item Only one robot wins an auction, and bids are updated for the remaining robots in subsequent auctions
	\item A receding-horizon framework optimizes the time to update the optimal future poses
	\item Cell degradation provides a simple means for autonomous patrol via a continuous-time Markov process
	\item Numerical examples of a large experimental space demonstrate both algorithms
\end{itemize}
\end{frame}

\section*{}
\subsection*{Experiments}

\begin{frame}
\frametitle{Experiments}
\begin{enumerate}[I]
\addtocounter{enumi}{3}
	\item Experiments
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Experimental Overview}

\begin{enumerate}
	\item Ground vehicle experiment for 2D occupancy grid mapping and autonomous exploration
	\item Aerial vehicle experiment in large 3D space resembling an office
	\item Aerial vehicle experiment in a complex 3D space
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Pioneer Ground Vehicle 2D Example}
\begin{itemize}
        	\item A pioneer ground vehicle explores a room composed of Styrofoam walls
	\item Experiments use ROS, a Microsoft Kinect depth sensor, and Vicon Tracker for pose information
\end{itemize}
\begin{figure}
	\centering
    	\begin{subfigure}[b]{0.28\textwidth}
        		\includegraphics[width=\textwidth]{pioneer.png}
        		\caption*{Pioneer}
    	\end{subfigure}    	
	\hspace*{0.04\textwidth}
	\begin{subfigure}[b]{0.28\textwidth}
        		\includegraphics[width=\textwidth]{test_setup_1.jpg}
        		\caption*{Bottom-left view}
    	\end{subfigure}
	\hspace*{0.03\textwidth}
	\begin{subfigure}[b]{0.28\textwidth}
        		\includegraphics[width=\textwidth]{test_setup_2.jpg}
        		\caption*{Bottom-right view}
    	\end{subfigure}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Ground Experiment Video}

	\centering{
		\includemedia[
  		width=0.720\textwidth,
  		height=0.576\textwidth,
  		activate=pageopen,
  		addresource=NRL_ground_test.mp4,
  		flashvars={source=NRL_ground_test.mp4}
		]{}{VPlayer.swf}
		}

\end{frame}

\begin{frame}
\frametitle{Ground Vehicle Experimental Results}

\begin{figure}
	\centering{
    	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_start.jpg}
        		\caption*{$t=0.0$min}
        		\label{fig:Experiment_ogm_t0}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_0p5min.jpg}
        		\caption*{$t=0.5$min}
        		\label{fig:Experiment_ogm_t0p5}
    	\end{subfigure}    
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1min.jpg}
        		\caption*{$t=1.0$min}
        		\label{fig:Experiment_ogm_t1}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_1p5min.jpg}
        		\caption*{$t=1.5$min}
        		\label{fig:Experiment_ogm_t1p5}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2min.jpg}
        		\caption*{$t=2.0$min}
        		\label{fig:Experiment_ogm_t2}
    	\end{subfigure}
		\vspace*{0.05\textwidth}

	}
	\vspace*{-1cm}
	\centering{
    	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_2p5min.jpg}
        		\caption*{$t=2.5$min}
        		\label{fig:Experiment_ogm_t2p5}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3min.jpg}
        		\caption*{$t=3.0$min}
        		\label{fig:Experiment_ogm_t3}
    	\end{subfigure}    
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_3p5min.jpg}
        		\caption*{$t=3.5$min}
        		\label{fig:Experiment_ogm_t3p5}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4min.jpg}
        		\caption*{$t=4.0$min}
        		\label{fig:Experiment_ogm_t4}
    	\end{subfigure}
	\begin{subfigure}[b]{0.19\textwidth}
        		\includegraphics[trim={13cm 1cm 13cm 0}, clip, width=\textwidth]{t_4p5min.jpg}
        		\caption*{$t=4.5$min}
        		\label{fig:Experiment_ogm_t4p5}
    	\end{subfigure}
	}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Aerial Vehicle in a Large Office-Like Environment}
\begin{itemize}
        	\item Hardware
	\begin{itemize}
		\item NVIDIA Jetson Ubuntu computer-on-module
		\item Asus Xtion IR color depth sensor
		\item Hokuyo LIDAR
		\item VectorNav IMU
	\end{itemize}
	\item Software
	\begin{itemize}
		\item Robot Operating System (ROS)
		\item Vicon Tracker (mocap\_vicon)
		\item Geometric Control: coordinate-free attitude control
	\end{itemize}
\end{itemize}
\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\linewidth}
    \centering\includegraphics[height=.5\linewidth]{quad_top.png}
    \caption*{Quadrotor Top}
  \end{subfigure}
  \begin{subfigure}[t]{.4\linewidth}
    \centering\includegraphics[height=.5\linewidth]{quad_bottom.png}
    \caption*{Quadrotor Bottom}
  \end{subfigure}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Aerial Vehicle in a Large Office-Like Environment}
\begin{itemize}
        	\item Environment Size: $10.8$ m $\times$ $11.8$ m $\times$ $1.65$ m
\end{itemize}
\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\linewidth}
    \centering\includegraphics[width=\linewidth]{experiment_north.jpg}
    \caption*{North Region}
  \end{subfigure}
  \begin{subfigure}[t]{.4\linewidth}
    \centering\includegraphics[width=\linewidth]{experiment_south.jpg}
    \caption*{South Region}
  \end{subfigure}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Large Flight Test Experiment Video}

	\centering{
		\includemedia[
  		width=0.720\textwidth,
  		height=0.576\textwidth,
  		activate=pageopen,
  		addresource=NRL_3D_large_space.mp4,
  		flashvars={source=NRL_3D_large_space.mp4}
		]{}{VPlayer.swf}
		}

\end{frame}

\begin{frame}
\frametitle{Aerial Vehicle in a Complex 3D Environment}
\begin{itemize}
        	\item Full 3D mapping and exploration
	\item Objects are not vertically-uniform
	\item Environment Size: $9.0$ m $\times$ $8.5$ m $\times$ $2.25$ m
\end{itemize}

\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\linewidth}
    \centering\includegraphics[height=.7\linewidth]{explore3D_space_NW.JPG}
    \caption*{Northwest View}
  \end{subfigure}
  \begin{subfigure}[t]{.4\linewidth}
    \centering\includegraphics[height=.7\linewidth]{explore3D_space_NE.JPG}
    \caption*{Northeast View}
  \end{subfigure}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Full 3D Aerial Experiment Video}

	\centering{
		\includemedia[
  		width=0.720\textwidth,
  		height=0.576\textwidth,
  		activate=pageopen,
  		addresource=explore3D_SEH.mp4,
  		flashvars={source=explore3D_SEH.mp4}
		]{}{VPlayer.swf}
		}

\end{frame}

\begin{frame}
\frametitle{Experiment Summary}
\begin{itemize}
        	\item Experimental environments in 2D and 3D are modeled with occupancy grids
	\item Future entropy prediction and collision-free regions use these occupancy probabilities
	\item A geometric controller is used during flight testing to follow the collision-free trajectory
	\item All algorithms are run in real-time without expensive post-processing 
\end{itemize}
\end{frame}

\section*{}
\subsection*{Conclusions}

\begin{frame}
\frametitle{Conclusions}
\begin{itemize}
        	\item Proposed an occupancy grid mapping technique that uses the \emph{exact probabilistic solution}
	\item Computational cost is reduced \emph{substantially} for \emph{real-time implementation} using probabilisitic properties and exploiting mathematical patterns
	\item Exact mapping concepts are extended to map uncertainty to accomplish \emph{autonomous exploration}
	\item Multiple vehicles can explore and patrol an uncertain environment using a \emph{bidding}-based approach
	\item These are demonstrated with numerous simulations and experiments
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Future Directions}
	\begin{itemize}
		\item Localization: use probabilistic map and other onboard sensors to localize, thereby completing the SLAM problem with the novel mapping approach
		\item Computational improvements: determine collision-free costs more efficiently than Dijkstra's cost map
		\item Decentralize the algorithms so processes may be distributed among multiple agents
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Publications on Dissertation Topics}
{\tiny 
\begin{itemize}
	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Multi-Vehicle Cooperative Exploration and Patrol of Uncertain 3D Environments'', in preparation.
	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Bidding-Based Autonomous Exploration of Cooperative Robots Building a 3D Probabilistic Occupancy Grid Map'', in preparation.
	\item E. Kaufman and T. Lee, ``Autonomous Aerial Exploration for Topological Mapping of Mars Environments'', \textit{2019 AIAA Guidance, Navigation, and Control Conference}, San Diego, CA, January 2019, submitted.
	\item E. Kaufman, K. Takami, Z. Ai, and T. Lee, ``Autonomous Quadrotor 3D Mapping and Exploration Using Exact Occupancy Probabilities,'' \textit{The Second IEEE International Conference on Robotic Computing}, pp. 49--55, Laguna Hills, CA, January 2018.
	\item E. Kaufman, K. Takami, T. Lee, and Z. Ai, ``Autonomous Exploration with Exact Inverse Sensor Models,'' \textit{Journal of Intelligent }\&\textit{ Robotic Systems}, 2017, doi: 10.1007/s10846-017-0710-7.
	\item E. Kaufman, T. Lee, and Z. Ai, ``Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping,'' \textit{IEEE International Conference on Simulation, Modeling, and Programming for Autonomous Robots}, pp. 246--251, San Francisco, CA, December 2016.
	\item E. Kaufman, T. Lee, Z. Ai, and I. S. Moskowitz, ``Bayesian Occupancy Grid Mapping via an Exact Inverse Sensor Model,'' \textit{Proceedings of the American Control Conference}, pp. 5709-5715, Boston, MA, July 2016.
\end{itemize}
}
\end{frame}

\begin{frame}
\frametitle{Publications on Data Association, Estimation, and Control}
{\tiny 
\begin{itemize}
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{The Journal of the Astronautical Sciences}, 63(1): pp. 60-80, 2016, doi: 10.1007/s40295-015-0082-9.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Minimum Uncertainty JPDA Filters and Coalescence Avoidance for Multiple Object Tracking,'' \textit{The Journal of the Astronautical Sciences}, 63(4): pp. 308-334, 2016, doi: 10.1007/s40295-016-0092-2.
	\item T. Wu, E. Kaufman, and T. Lee, ``Globally Asymptotically Stable Attitude Observer on SO(3),'' \textit{Proceedings of the 54th IEEE Conference on Decision and Control}, pp. 2164-2168, Osaka, Japan, December 2015.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Nonlinear Observability Measure for Relative Orbit Determination with Angles-Only Measurements,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting}, AAS 15-451, Williamsburg, VA, January 2015.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Minimum Uncertainty JPDA Filter and Coalescence Avoidance Performance Evaluations,'' \textit{Proceedings of the 25th AAS/AIAA Space Flight Mechanics Meeting, Williamsburg}, AAS 15-432, VA, January 2015.
	\item E. Kaufman, K. Caldwell, D. Lee, and T. Lee, ``Design and Development of a Free-Floating Hexrotor UAV for 6-DOF Maneuvers,'' \textit{Proceedings of the IEEE Aerospace Conference}, ASC 14-2527, March 2014, Big Sky, MT.
	\item E. Kaufman, T. A. Lovell, and T. Lee, ``Optimal Joint Probabilistic Data Association Filter Avoiding Coalescence in Close Proximity,'' \textit{Proceedings of the European Control Conference}, pp. 2709-2714, Strasbourg, June 2014.
\end{itemize}
}
\end{frame}

\begin{frame}
\frametitle{Thanks}
\begin{itemize}
        	\item Special thanks to the US Naval Research Lab for support under project ``Intelligent Microflyer''
	\item Questions
\end{itemize}
\end{frame}

\end{document}